---
title: "compile_datasheets_HR_2023 "
author: "Helen and Devin"
date: "2022-08-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message = FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)
library(stringr)

```



### Load and Clean Data   

**Fitness Plant Data**  

Recipient IDs to double-check:
- HR_067 (HR_067B)

- Add in Donor information
```{r message = FALSE}
HR_Fitness <- read_csv(here::here("data_sheets", "2023_year2_data", "fitness_plants", "HR_fitnessplants_23.csv")) %>% 
  mutate(Recipient = case_when(Recipient %in% c("HR_067", "HR_067B") ~ "HR_067",
                               TRUE ~ Recipient))  #rename - letters were added to mark position

HR_donors <- read_csv(here::here("data_sheets", "2023_year2_data", "planting_locations", "HR_donor_info.csv")) 

#Fix Recipient and Donor names
HR_donors2 <- HR_donors %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("HR_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("HR_", Donor_num))) %>% 
  select(-Recipient_num, -Donor_num)

#Multiple matches, merge only unique pairs for Donor
HR_Fitness_2 <- merge(HR_Fitness, unique(HR_donors2[c("Gen", "Donor", "Recipient")]), by = c("Recipient", "Gen"), all.x = TRUE) %>% 
  mutate(SegPos = paste(Segment, position, sep = "")) %>% 
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Fit_plant) %>% 
  arrange(Sequence, SegPos)

#Save as V2 Fitness Plant Sheet
write_csv(HR_Fitness_2, here::here("data_sheets", "2023_year2_data", "fitness_plants", "HR_fitnessplants_23_V2.csv"), na = "")


```

<br>  

### Trait Data  

Trait data gathered and processed from the 2022-2023 field season is loaded and organized below.  


#### Flower Check & Phenology  
-- Organize flowering data
-- Add in Sow & Plant dates

#########
Crucial Step!!  
######### 
In the csv file, select all data with dates (visit dates, leaf sampling, and flower photos). Right click --> format cells. Select the date category --> YYYY-MM-DD. Save the csv and close it. This will standardize all dates to the same format and make working with them in R (and lubridate) much easier!  

```{r message = FALSE}
HR_flr_df <- read_csv(file = here::here( "data_sheets", "2023_year2_data", "flowering_checks", "HR_flowercheck_data_2023.csv"), na = c("", "NA"))  %>% 
  mutate(Recipient = case_when(Recipient %in% c("HR_067", "HR_067B") ~ "HR_067",
                               TRUE ~ Recipient))  #rename - letters were added to mark position


HR_flr2 <- HR_flr_df %>% 
  mutate(SegPos = paste(Segment, side, sep = "")) %>% 
  select(Sequence, Transect, SegPos, everything(), -c(Segment, side, FD_06, FD_07)) %>% 
  arrange(Sequence, Transect, SegPos) %>% 
  mutate(Recipient = case_when(Recipient %in% c("HR_067", "HR_067B") ~ "HR_067",
                               TRUE ~ Recipient)) #rename- letters were added to mark position

#Check for mismatched fitness plants
HR_flr_test <- HR_flr2 %>% 
  filter(!is.na(Fit_plant)) %>% 
  mutate(FitMatch = case_when(Fit_plant == Plant_ID ~ TRUE,
                              Fit_plant != Plant_ID ~ FALSE))

## Fix Dates!

#Check obs first using unique(HR_flr3$FD_01) and make corrections

HR_flr3 <- HR_flr2 %>%
  mutate_at(.vars = vars(FD_01:FD_05), ~as_date(.)) #the warning is because of the 4 "G"s included for Gopher in FD_03 and FD_04

```

Add sowing/planting date  
```{r}
#Sow Date - first date (of two or one) on which seeds were sown in the lab
#Plant Date - median date on which seed plugs were planted in the field

HR_flr4 <- HR_flr3 %>% 
  mutate(Sow_Date = as_date("2022-11-10"),
         Plant_Date = as_date("2022-11-13"))

```


Merge in Fitness plant Data  
- Donor

```{r}
#Check to make sure fitness plants line up
HR_flr5 <- merge(HR_flr4, HR_Fitness_2, by = c("Sequence", "Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE) %>%
  arrange(Sequence, Transect, SegPos, Plant_ID) %>% 
  mutate(F_plant = case_when(!is.na(Fit_plant.x) ~ TRUE,
                               is.na(Fit_plant.x) ~ FALSE)) %>% 
  select(-c(Fit_plant.x, Fit_plant.y)) #Drop extra fitplant row after checking that they all match - retain multiple Donors per segment
  
  
#####  
#Check to make sure fitness plants line up
  #HR_flr_test2 <- HR_flr5 %>% 
  #  filter(!is.na(Fit_plant.x)) %>%
  #  mutate(FMatch = case_when(Fit_plant.x == Fit_plant.y ~ TRUE,
  #                          Fit_plant.x != Fit_plant.y ~ FALSE))

```


**Phenology Calculations**    

Survivorship calculated later

```{r}
#Same as last year

HR_flowering <- HR_flr5 %>% 
  mutate(FFD = do.call(pmin, c(select(., starts_with('FD_')), na.rm = TRUE)), #First flowering date
         days_sow2flower = as.numeric(FFD - Sow_Date), #days from lab sowing to flowering
         days_plant2flower = as.numeric(FFD - Plant_Date), #days from field planting to flowering
         LFD = do.call(pmax, c(select(., starts_with('FD_')), na.rm = TRUE)), #last flowering date
         fl_duration = as.numeric(LFD - FFD)) %>%  #Flowering duration
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, Sow_Date, Plant_Date, FFD, days_sow2flower, days_plant2flower, LFD, fl_duration, FD_01:FD_05, everything())


#TEMP FILE
write_csv(HR_flowering, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "HR_phenology_2023.csv"), na = "")

```

<br>  



#### Fruit and Seed counts    

Fruit & seed counts were recorded on the same spreadsheet, unlike last year. Loaded here and checked to make sure all ID info lines up with the Fitness data sheet.

```{r message = FALSE}
#Load Fruit & Seeds count data

HR_fruit_seeds <- read_csv(here::here("data_sheets", "2023_year2_data", "fruit_seed_counts", "HR_fruits_seeds_2023 - HR_fruit_seed_counts.csv"), skip = 1) %>% 
  rename(SegPos = Seg) %>%
  mutate(Recipient = case_when(Recipient %in% c("HR_067", "HR_067B") ~ "HR_067",
                               TRUE ~ Recipient))  #rename - letters were added to mark position


# Remove rows where no fruits collected  (optional)
# HR_fruit_seeds_2 <- HR_fruit_seeds %>% 
#   filter(total_fruits != 0)


##Check Fitness plants
HR_merge_test <- merge(HR_Fitness_2, HR_fruit_seeds, by = c("Transect", "SegPos", "Gen", "Recipient", "Fit_plant"), all.y = TRUE) 


HR_fruit_seeds_3 <- merge(HR_Fitness_2, HR_fruit_seeds, by = c("Transect", "SegPos", "Gen", "Recipient", "Fit_plant"), all.x = TRUE) %>% 
  select(-Sequence.y) %>% 
  rename(Sequence = Sequence.x) %>% 
  rename(Plant_ID = Fit_plant)

```

<br>  

#### Skeleton Weights  

Load and merge-in these data sheets
```{r message = FALSE}

skel_weights_HR <- read_csv(here::here("data_sheets", "2023_year2_data",  "skeleton_data", "Skeleton Data 2023 - HR Skeletons.csv")) %>% 
  rename(position = side, Gen = 4, Fitness = 7, skel_weight_mg = 8, complete_incomplete = 9, selected = 10) %>% 
  arrange(Transect, Segment, position) %>% 
  mutate(SegPos = paste(Segment, position, sep = "")) %>% 
  select(Transect, SegPos, Gen,  Recipient, Plant_ID, Fitness, skel_weight_mg) %>%  #complete/selected not included
  mutate(Recipient = case_when(Recipient %in% c("HR_067", "HR_067B") ~ "HR_067",
                               TRUE ~ Recipient)) #rename- letters were added to mark position


skel_weights_HR_2 <- skel_weights_HR %>% 
  mutate(Recipient = case_when(Recipient %in% c("HR_067", "HR_067B") ~ "HR_067",
                               TRUE ~ Recipient)) #rename- letters were added to mark position

#Check that IDs line up
skelcheck <- merge(HR_Fitness_2, skel_weights_HR_2, by = c("Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE) %>% 
  mutate(FTRUE = case_when(Fit_plant == Plant_ID ~ TRUE,
                           TRUE ~ FALSE))


```



**TEMP MERGED DATA FRAME**  
Merge trait data processed thus far

```{r}

HR_merged_1 <- merge(HR_flowering, skel_weights_HR, by = c("Transect", "SegPos", "Gen", "Recipient", "Plant_ID"), all.x = TRUE) %>% 
  select(-c(Skeleton:Fitness)) %>% 
  arrange(Transect, Sequence, SegPos)

HR_merged_2 <- merge(HR_merged_1, HR_fruit_seeds_3, by = c("Sequence", "Transect", "SegPos", "Gen", "Recipient", "Donor", "Plant_ID"), all.x = TRUE) %>% 
  select(-c(FD_01:FD_05, N_CLOSED_fruit_D1:notes)) %>% 
  arrange(Transect, Sequence, SegPos) %>% 
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, everything())
```

<br>  


#### Seed Weights  
[[UPDATED]]

```{r eval = FALSE}

HR_seedmass_1 <- read_csv(here::here("data_sheets", "seed_mass", "HR_Seed_Mass_2023.csv"), skip = 1) %>% 
  rename(SegPos = 3, Plant_ID = 7, mean_seed_mass_mg = 8) %>% 
  select(Transect, SegPos, Gen,  Recipient, Plant_ID, mean_seed_mass_mg)
  
##Merge with fruit/seed count data frame

HR_merged_3 <- merge(HR_merged_2, HR_seedmass_1, by = c("Transect", "SegPos", "Recipient", "Plant_ID", "Gen"), all.x = TRUE)
#all.x means that any fitness plant x ID combos that don't match up won't be included.

```

<br>  




#### Corolla measurements  
[[UPDATED]]

```{r eval = FALSE}
corollas_HR <- read_csv(here::here("data_sheets", "corolla_measures", "HR Corolla Diameters 2023.csv"), skip = 1) %>% 
  rename(Transect = 1, Recipient = 2, Gen = 3, Plant_ID = 5, SegPos = 6, corolla_diam_mm = 7, photo_date = 8) %>% 
  select(Transect, Recipient, Gen, Plant_ID, SegPos, corolla_diam_mm, photo_date) %>%  #complete/selected not included
  mutate(Recipient = case_when(Recipient %in% c("HR_067", "HR_067B") ~ "HR_067",
                               TRUE ~ Recipient)) #rename- letters were added to mark position

corollas_HR2 <- corollas_HR %>% 
  filter(!is.na(corolla_diam_mm)) %>% 
  mutate(photo_date = as_date(photo_date)) %>% #correct missing dates manually
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% #Fix Recipient names
  mutate(Recipient = as.character(str_c("HR_", Recipient_num))) %>% 
  select(-Recipient_num, -Plant_ID) %>% 
  arrange(Transect, Recipient, photo_date) %>% 
  distinct(Transect, Recipient, .keep_all = TRUE) #Keep only the earliest photo date data for duplicates
#remove plant ID bc we used NF plants occasionally to obtain corolla diameter


## Merge with main data 
HR_merged_4 <- merge(HR_merged_3, corollas_HR2, by = c("Transect", "Recipient", "Gen", "SegPos"), all.x = TRUE)
#did not include plant ID in the merge bc we got data for NF plants
```

<br>  


#### Leaf Area & Weight (SLA)  
[[UPDATED]]

```{r eval = FALSE}
leaves_HR <- read_csv(here::here("data_sheets", "leaf_data", "HR_Leaf_Area_Weight_2023.csv"), skip = 1) %>% 
  rename(Transect = 1, Gen = 2, Donor= 3, Recipient = 4, SegPos = 5, Plant_ID = 6, leaf_area_mm2 = 7, leaf_dryweight_mg = 10) %>% 
  select(Transect, Recipient, Gen, Donor, SegPos, leaf_area_mm2, leaf_dryweight_mg) %>% #didn't include Plant_ID because corolla area will could be from either fitness or non-fitness plants 
  filter(!is.na(leaf_area_mm2)) %>% #exlcude NA leaf areas?
  mutate(leaf_area_mm2 = as.numeric(leaf_area_mm2))


#Chec for duplicates
leaves_HR_dup_check <- leaves_HR %>% 
  mutate(duped1 = duplicated(select(leaves_HR, c(Transect, Recipient, Gen, Donor, SegPos))),
         duped2 = duplicated(select(leaves_HR, c(Transect, Recipient, Gen, Donor, SegPos)), fromLast = TRUE),
         duped = case_when(duped1|duped2 == TRUE ~ TRUE,
                           TRUE ~ FALSE)) %>% #highlight all duplicates
  select(-c(duped1, duped2))
# No "TRUE"s found, so no duplicates, no need to remove them

#Merge in to main data 

HR_merged_5 <- merge(HR_merged_4, leaves_HR, by = c("Transect", "Recipient", "Donor", "Gen", "SegPos"), all.x = TRUE) #replicates leaf area & weight for every plant of a transect recipient combo
#did not include plant ID in the merge bc we got data for NF plants

#Create a column called SLA in the merged dataframe
HR_merged_6 <- HR_merged_5 %>%
  mutate(SLA = leaf_area_mm2 / leaf_dryweight_mg)
```

<br>  



#### d13C data (WUE)  
[[UPDATED]]

```{r eval = FALSE}
HR_d13C_2023 <- read_csv(here::here("data_sheets", "isotope_data", "HR_d13C_2023.csv")) %>% 
  select(1,2,3,5,6,9) %>% 
  rename(sample_ID = 1, d13C = 2, totalC_ug = 3, sample_weight_mg = 6)
#Save as UTF-8 file from excel to avoid import error

#unravel sample ID
HR_d13C_2023_2 <- HR_d13C_2023 %>% 
mutate(Recipient = gsub('-', '_', substr(sample_ID, 1, 6)), 
       Transect = as.numeric(str_remove(substr(sample_ID, 11, 13), "^0+")),
       Gen = as.character(substr(sample_ID, 8, 9)),
       SegPos = gsub('-', '', substr(sample_ID, 15, 17)),
        Plant_ID = gsub('-', '', substr(sample_ID, 18, 19))) %>% 
 select(Transect, Gen, Recipient, SegPos, Plant_ID, d13C, sample_weight_mg, -Plant_ID)
  
  
#Merge
HR_merged_7 <- merge(HR_merged_6, HR_d13C_2023_2, by = c("Recipient", "Transect", "Gen", "SegPos"), all.x = TRUE)
#Merge by Recipient and Transect only: genotype-wide estimates of d13C since some plants were NF



# Subset to exclude rows where d13C is NA
testdf <- HR_merged_7 %>% #one non-fitness plant skeleton was used HR_10_T1_G1_P4_1L
  filter(F_plant == "TRUE") %>% #603 vs 604
  filter(!is.na(d13C))   # Exclude rows where d13C is NA

```

**Error checks**  

Checking a variety of IDs and trait data to make sure everything lines up.  

```{r eval = FALSE}
##Check for mismatched (fitness) plants

HR_Echeck1 <- HR_merged_7 %>% 
  filter(F_plant == FALSE & total_fruits != 0) #do the same for seed count, using `>` and `!=`

HR_Echeck2 <- HR_merged_7 %>% 
  filter(F_plant == TRUE) %>% 
  filter(ALL_closed_fruit == 0 & total_seeds > 0)

##NA/NaN/Info values for mean seeds/est seed prod
HR_Echeck3 <- HR_merged_7 %>% 
  filter(F_plant == TRUE & is.na(as.numeric(avg_seeds_per_fruit)))

HR_Echeck4 <- HR_merged_7 %>% 
  filter(F_plant == TRUE & !is.finite(avg_seeds_per_fruit))


## Check for NA seed counts for fitness plants (when fruit production > 0)
#Will be empty when seed_ct NAs are replaced with 0
HR_Echeck5 <- HR_merged_7 %>% 
  filter(F_plant == TRUE & is.na(total_seeds))
##NAs present here before setting seed_ct NAs to 0 were from fitness plants that had 0 closed fruit


#Seed mass
HR_Echeck6 <- HR_merged_7 %>% 
  filter(F_plant == FALSE & !is.na(mean_seed_mass_mg))


```

<br>  


### Trait calculations 

Some trait calculations performed in `Fitness_and_EDA.Rmd.  

Fruits were only collected from fitness plants, so only survival to flowering is calcualted for all plants (fitness and non-fitness).

UPDATE with latest merged df...

```{r}

#Survival measures
#Added var for closed fruit produced
#renamed "survival" to seed production
#
HR_merged_8 <- HR_merged_7 %>% 
  mutate(surv_to_flower = case_when(!is.na(FFD) ~ 1, #plants that survived to flower
                                  is.na(FFD) ~ 0),
         surv_to_fruitprod = case_when(F_plant == TRUE & total_fruits > 0 ~ 1, #plants that survived to produce fruit
                                       F_plant == TRUE & total_fruits == 0 ~ 0),
         prod_closed_fruits = case_when(F_plant == TRUE & ALL_closed_fruit > 0 ~ 1,
                                            F_plant == TRUE & ALL_closed_fruit == 0 ~ 0),
         prod_filled_seeds = case_when(F_plant == TRUE & ALL_filled_seeds > 0 ~ 1, #plants produced filled seeds
                                      F_plant == TRUE & ALL_filled_seeds == 0 ~ 0)) #else = 0

#NOTE
#Adjusting fecundity/fitness:
#The issue of how to handle 0s vs missing data in order to accurately calculate fitness has led us to apply the following corrections:
# Estimated fecundity (seeds per fruit x total fruits) should be set to *0* when closed fruits >= 1 & filled seeds == 0; when closed fruits == open fruits == 0
#Set fecundity to NA when closed = 0 & open >= 1

#Mean seeds per fruit left as-is (NA when no closed fruits collected)
         

HR_merged_9 <- HR_merged_8 %>% 
  rename(mean_seeds_per_fruit = avg_seeds_per_fruit) %>%
  mutate(across(c(est_fecundity, mean_seeds_per_fruit), ~ as.numeric(na_if(., NaN)))) %>% 
#Converts to NA if any of the values are "NaN"
  mutate(mean_seeds_per_fruit = case_when(
          F_plant == TRUE & ALL_closed_fruit == 0 ~ NA, # NA when 0 closed fruits
          TRUE ~ mean_seeds_per_fruit)) %>%  #else, original value
          mutate(est_fecundity = case_when(
          F_plant == TRUE & ALL_closed_fruit == 0 & ALL_open_fruit == 0 ~ 0,
          F_plant == TRUE & ALL_closed_fruit == 0 & ALL_open_fruit >= 1 ~ NA,
          F_plant == TRUE & !is.na(est_fecundity) ~ est_fecundity))


```

<br>  



### Master sheets (2023)  

Full and summarized data sheets for all data collected in the 2022-2023 field season. 
These are merged with the year 1 data sheets elsewhere.  

```{r}
## 'Long' version with all data - one row per plant ID
#Sort by transect and ID variables, then by Fitness plants and those that flowered...

######
######

HR_merged_full_2023 <- HR_merged_9 %>% 
  filter(F_plant == TRUE)
##NOTES - there are still fitness plants that never flowered... keep as 0s
#fitness plants only

write_csv(HR_merged_full_2023, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "HR_merged_full_2023.csv"), na = "")


######
######


HR_mastersheet_v1_2023 <- HR_merged_9 %>% 
  mutate(Year = 2023) %>% 
  select(Year, Gen, Transect, Sequence, Donor, Recipient, Plant_ID, F_plant, everything()) %>% 
  arrange(Transect, Sequence, Recipient, Plant_ID)

write_csv(HR_mastersheet_v1_2023, here::here("data_sheets", "compiled_sheets", "HR_mastersheet_full_2023.csv"))
#Renamed to full - has records of all plants (including non-fitness)



#
## Fitness Plants Only
#

HR_mastersheet_F_mains_2023 <-  HR_mastersheet_v1_2023 %>% 
  filter(F_plant == TRUE) #Only main fitness plants

##Note: remaining NA values for fitness measures (mean seeds per fruit; est seed production) are the results of 0 closed fruit counts. Treat these as 0 fitness or as NAs to remove???


write_csv(HR_mastersheet_F_mains_2023, here::here("data_sheets", "compiled_sheets", "HR_mastersheet_Fitness-mains_2023.csv"))

```



