---
title: "compile_datasheets_AC_2023"
author: "Helen Payne and Devin Gamble"
date: "2024-23-05"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: default
---

```{r}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)
library(stringr)

```

###ISSUE####
Genotype Segments that had 0 survivors and NO fitness plant selected are EXCLUDED as of now... These should probably be included but with an assignment of 0 fitness


### Load and Clean Data   

Recipient IDs to double-check:
- AC_159 (AC_159A/B)
- IDs not planted

**Load Fitness Plant Data**  

```{r}
AC_Fitness <- read_csv(here::here("data_sheets", "2023_year2_data", "fitness_plants", "AC_fitness_23.csv")) %>% 
  slice(-639) %>% #removing the random seeds planted in Transect 9, seg 34 (left)
  mutate(Recipient = case_when(Recipient %in% c("AC_159", "AC_159A", "AC_159B") ~ "AC_159",
                               TRUE ~ Recipient)) %>%  #rename 159 - letters were added to mark position
  slice(-423) #AC_213 in Tr 6 seg 33L was never planted - remove entirely

AC_donors <- read_csv(here::here("data_sheets", "2023_year2_data", "planting_locations", "AC_donor_info.csv")) %>% 
  select(-position) 

#Fix Recipient and Donor names
AC_donors2 <- AC_donors %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("AC_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("AC_", Donor_num))) %>% 
  select(-Recipient_num, -Donor_num)



#Multiple matches, merge only unique pairs for Donor
AC_Fitness_2 <- merge(AC_Fitness, unique(AC_donors2[c("Gen", "Donor", "Recipient")]), by = c("Recipient", "Gen"), all.x = TRUE) %>% 
  mutate(SegPos = paste(Segment, position, sep = "")) %>% 
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Fit_plant) %>% 
  arrange(Sequence, SegPos)

#Save as V2 Fitness Plant Sheet
write_csv(AC_Fitness_2, here::here("data_sheets", "2023_year2_data", "fitness_plants", "AC_fitnessplants_23_V2.csv"), na = "")


```



### Trait Data  

Trait data gathered and processed from the 2022-2023 field season is loaded and organized below.  


#### Flower Check & Phenology  
-- Organize flowering data
-- Add in Sow & Plant dates

#########
Crucial Step!!  
######### 
In the csv file, select all data with dates (visit dates, leaf sampling, and flower photos). Right click --> format cells. Select the date category --> YYYY-MM-DD. Save the csv and close it. This will standardize all dates to the same format and make working with them in R (and lubridate) much easier!  

```{r}

AC_flr_df <- read_csv(file = here::here( "data_sheets", "2023_year2_data", "flowering_checks", "AC_flowercheck_data_2023.csv"), na = c("", "NA"))


AC_flr2 <- AC_flr_df %>% 
  mutate(SegPos = paste(Segment, side, sep = "")) %>% 
  select(Sequence, Transect, SegPos, everything(), -c(Segment, side, FD_06, FD_07)) %>% 
  slice(-c(1355, 1600, 2111:2115, 3191:3195)) %>% #Remove rows not planted
  arrange(Sequence, Transect, SegPos) %>% 
  mutate(Recipient = case_when(Recipient %in% c("AC_159", "AC_159A", "AC_159B") ~ "AC_159",
                               TRUE ~ Recipient)) #rename 159 - letters were added to mark position
  

#Check for mismatched fitness plants
AC_flr_test <- AC_flr2 %>% 
  filter(!is.na(Fit_plant)) %>% 
  mutate(FitMatch = case_when(Fit_plant == Plant_ID ~ TRUE,
                              Fit_plant != Plant_ID ~ FALSE))

## Fix Dates!

#Check obs first using unique(AC_flr3$FD_01) and make corrections

AC_flr3 <- AC_flr2 %>%
  mutate_at(.vars = vars(FD_01:FD_05), ~as_date(.)) %>% 
  mutate(FD_05 = as_date(case_when(FD_05 == "2023-06-07" ~ "2023-06-08",
                           FD_05 == "2023-06-08" ~ "2023-06-08",
                           FD_05 == "2023-06-09" ~ "2023-06-09",
                           TRUE ~ as.character(NA)))) #Some date typos - one day early

```

Add sowing/planting date  
```{r}
#Sow Date - first date (of two or one) on which seeds were sown in the lab
#Plant Date - median date on which seed plugs were planted in the field

AC_flr4 <- AC_flr3 %>% 
  mutate(Sow_Date = as_date("2022-10-17"),
         Plant_Date = as_date("2022-10-21"))

```


Merge in Fitness plant Data  
- Donor
```{r}
#Check to make sure fitness plants line up
AC_flr5 <- merge(AC_flr4, AC_Fitness_2, by = c("Sequence", "Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE) %>%
  arrange(Sequence, Transect, SegPos, Plant_ID) %>% 
  mutate(F_plant = case_when(!is.na(Fit_plant.x) ~ TRUE,
                               is.na(Fit_plant.x) ~ FALSE)) %>%
 select(-c(Fit_plant.x, Fit_plant.y)) #Drop extra fitplant row after checking that they all match - retain multiple Donors per segment
  
  
#####  
#Check to make sure fitness plants line up, compare this by commenting out the "select" above, and see if this df now matches the size of the test file. They both have 546 variables, so we are good!
  #filter(!is.na(Fit_plant.x)) %>%
#mutate(FMatch = case_when(Fit_plant.x == Fit_plant.y ~ TRUE,
                            #Fit_plant.x != Fit_plant.y ~ FALSE))

```


**Phenology Calculations**  

```{r}
#Same as last year

AC_flowering <- AC_flr5 %>% 
  mutate(FFD = do.call(pmin, c(select(., starts_with('FD_')), na.rm = TRUE)), #First flowering date
         days_sow2flower = as.numeric(FFD - Sow_Date), #days from lab sowing to flowering
         days_plant2flower = as.numeric(FFD - Plant_Date), #days from field planting to flowering
         LFD = do.call(pmax, c(select(., starts_with('FD_')), na.rm = TRUE)), #last flowering date
         fl_duration = as.numeric(LFD - FFD)) %>%  #Flowering duration
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, Sow_Date, Plant_Date, FFD, days_sow2flower, days_plant2flower, LFD, fl_duration, FD_01:FD_05, everything())



#TEMP FILE - delete later
write_csv(AC_flowering, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "AC_phenology_2023.csv"), na = "")

```

<br>  


#### Fruit and Seed counts  

Fruit & seed counts were recorded on the same spreadsheet, unlike last year. Loaded here and checked to make sure all ID info lines up with the Fitness data sheet.


```{r}
#Load Fruit & Seeds count data

AC_fruit_seeds <- read_csv(here::here("data_sheets", "2023_year2_data", "fruit_seed_counts", "AC_fruits_seeds_2023 - AC_fruit_seed_counts.csv"), skip = 1) %>% 
  rename(SegPos = Seg)


# #Remove rows where no fruits collected (optional)
# AC_fruit_seeds_2 <- AC_fruit_seeds %>% 
#   filter(total_fruits != 0)


##Check Fitness plants
AC_merge_test <- merge(AC_Fitness_2, AC_fruit_seeds, by = c("Transect", "SegPos", "Gen", "Recipient"), all.y = TRUE) %>% 
  mutate(FITTRUE = case_when(Fit_plant.x == Fit_plant.y ~ TRUE,
                             Fit_plant.x != Fit_plant.y ~ FALSE))


AC_fruit_seeds_2 <- merge(AC_Fitness_2, AC_fruit_seeds, by = c("Transect", "SegPos", "Gen", "Recipient", "Fit_plant"), all.x = TRUE) %>% 
  select(-Sequence.y) %>% 
  rename(Sequence = Sequence.x,
         Plant_ID = Fit_plant) #Rename Plant ID for merging below AFTER double-checking fitness plants
```

Note: Some Segments not represented by any Fitness plants (no plants survived) -- Fecundity values are currently NA but will be adjusted to 0 later. 

<br>  



#### Skeleton Weights  
```{r}
skel_weights_AC <- read_csv(here::here("data_sheets", "2023_year2_data",  "skeleton_data", "Skeleton Data 2023 - AC Skeletons.csv")) %>% 
  rename(position = side, Gen = 4, Fitness = 7, skel_dryweight_mg = 8, complete_incomplete = 9, selected = 10) %>% 
  arrange(Transect, Segment, position) %>% 
  mutate(SegPos = paste(Segment, position, sep = "")) %>% 
  select(Transect, SegPos, Gen,  Recipient, Plant_ID, Fitness, skel_dryweight_mg)  %>%  #complete/selected not included
  mutate(Recipient = case_when(Recipient %in% c("AC_159", "AC_159A", "AC_159B") ~ "AC_159",
                               TRUE ~ Recipient)) #rename 159 - letters were added to mark

#create a dataframe with skeleton weights for each segment
skel_weights_AC_2 <- skel_weights_AC %>% 
  rename(skel_dryweight_mg_SEG = 7) %>%
  filter(Fitness == "y") %>% #remove the NF plant info from the segment data so that it does not duplicate the maternal IDs 
  select(-Plant_ID)

#Check that IDs line up
skelcheck <- merge(AC_Fitness_2, skel_weights_AC_2, by = c("Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE)
```


<br>  

**TEMP MERGED DATA FRAME**  

Merge trait data processed thus far

**NOTE** SHOULD THOSE COLUMNS BE PRESENT/ABSENT IN THE FINAL DATA SHEET??

```{r}
#merging the skeleton data for each individual plant
AC_merged_1 <- merge(AC_flowering, skel_weights_AC, by = c("Transect", "SegPos", "Gen", "Recipient", "Plant_ID"), all.x = TRUE)  %>% select(-c(Fitness, FD_01:FD_05, LeafD, FLR_P, Skeleton, Notes)) %>% arrange(Transect, Sequence, SegPos)

#merging the skeleton data for the entire segment
AC_merged_1.5 <- merge(AC_merged_1, skel_weights_AC_2, by = c("Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE) %>% select(-Fitness) %>% arrange(Transect, Sequence, SegPos)

####<-left off here
AC_merged_2 <- merge(AC_merged_1.5, AC_fruit_seeds_2, by = c("Sequence", "Transect", "SegPos", "Gen", "Recipient", "Donor", "Plant_ID"), all.x = TRUE) %>% select(-c(N_CLOSED_fruit_D1:N_total_seeds_D5, Sym))  %>% arrange(Transect, Sequence, SegPos) %>% 
  select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, everything(), -28)

```


#### Seed Weights  
[[UPDATED THE DATASHEET]]  

Helen's Update: mean seed mass_mg was already calculated, we just want to add that to the merged dataframes
```{r}

AC_seedmass_1 <- read_csv(here::here("data_sheets", "2023_year2_data", "seed_mass", "AC_Seed_Mass_2023.csv"), skip = 1) %>% 
  rename(SegPos = 3, Plant_ID = 7, mean_seed_mass_mg = 8) %>% 
  select(Transect, SegPos, Gen,  Recipient, Plant_ID, mean_seed_mass_mg)
  
##Merge with fruit/seed count data frame

AC_merged_3 <- merge(AC_merged_2, AC_seedmass_1, by = c("Transect", "SegPos", "Recipient", "Plant_ID", "Gen"), all.x = TRUE)
#all.x means that any fitness plant x ID combos that don't match up won't be included.

```

#### Corolla measurements  
[[UPDATED]]  

```{r}
corollas_AC <- read_csv(here::here("data_sheets", "2023_year2_data", "corolla_leaf_data", "AC Corolla Diameters 2023.csv"), skip = 1) %>% 
  rename(Transect = 1, Recipient = 2, Gen = 3, Plant_ID = 5, SegPos = 6, corolla_diam_mm = 7, photo_date = 8) %>% 
  select(Transect, Recipient, Gen, Plant_ID, SegPos, corolla_diam_mm, photo_date)

corollas_AC_1.5 <- corollas_AC %>% 
  filter(!is.na(corolla_diam_mm)) %>% 
  mutate(photo_date = as_date(photo_date)) %>% #correct missing dates manually
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% #Fix Recipient names
  mutate(Recipient = as.character(str_c("AC_", Recipient_num))) %>% 
  select(-Recipient_num) %>% 
  arrange(Transect, Recipient, photo_date) %>% 
  distinct(Transect, Recipient, .keep_all = TRUE) %>% #Keep only the earliest photo date data for duplicates
  mutate(Recipient = case_when(Recipient %in% c("AC_159", "AC_159A", "AC_159B") ~ "AC_159",
                               TRUE ~ Recipient)) #rename 159 - letters were added to mark

## Merge with main data 
AC_merged_4 <- merge(AC_merged_3, corollas_AC_1.5, by = c("Transect", "Recipient", "Gen", "SegPos", "Plant_ID"), all.x = TRUE)

#create a dataframe with corolla areas for each segment
corollas_AC_2 <- corollas_AC %>% 
  filter(!is.na(corolla_diam_mm)) %>% 
  mutate(photo_date = as_date(photo_date)) %>% #correct missing dates manually
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% #Fix Recipient names
  mutate(Recipient = as.character(str_c("AC_", Recipient_num))) %>% 
  select(-Recipient_num, -Plant_ID) %>% 
  arrange(Transect, Recipient, photo_date) %>% 
  distinct(Transect, Recipient, .keep_all = TRUE) %>% 
  rename (corolla_diam_mm_SEG = 5, photo_date_SEG = 6) #Keep only the earliest photo date data for duplicates


AC_merged_4.5 <- merge(AC_merged_4, corollas_AC_2, by = c("Transect", "Recipient", "Gen", "SegPos"), all.x = TRUE)
```


<br>  


#### Leaf Area & Weight  
[[UPDATED]]

```{r}
leaves_AC <- read_csv(here::here("data_sheets", "2023_year2_data", "corolla_leaf_data", "AC_Leaf_Area_Weight_2023.csv"), skip = 1) %>% 
  rename(Transect = 1, Gen = 2, Donor= 3, Recipient = 4, SegPos = 5, Plant_ID = 6, leaf_area_mm2 = 7, leaf_dryweight_mg = 10) %>% 
  select(Transect, Recipient, Gen, Donor, SegPos, Plant_ID, leaf_area_mm2, leaf_dryweight_mg) %>%
  filter(!is.na(leaf_area_mm2)) %>% #exlcude NA leaf areas?
  mutate(leaf_area_mm2 = as.numeric(leaf_area_mm2))


#Chec for duplicates
leaves_AC_dup_check <- leaves_AC %>% 
  mutate(duped1 = duplicated(select(leaves_AC, c(Transect, Recipient, Gen, Donor, SegPos))),
         duped2 = duplicated(select(leaves_AC, c(Transect, Recipient, Gen, Donor, SegPos)), fromLast = TRUE),
         duped = case_when(duped1|duped2 == TRUE ~ TRUE,
                           TRUE ~ FALSE)) %>% #highlight all duplicates
  select(-c(duped1, duped2))
# No "TRUE"s found, so no duplicates, no need to remove them

#Merge in to main data 
AC_merged_5 <- merge(AC_merged_4.5, leaves_AC, by = c("Transect", "Recipient", "Donor", "Gen", "SegPos", "Plant_ID"), all.x = TRUE) #replicates leaf area & weight for every plant of a transect recipient combo

#create a dataframe with skeleton weights for each segment
leaves_AC_2 <- leaves_AC %>%
  rename(leaf_area_mm2_SEG = 7, leaf_dryweight_mg_SEG = 8) %>%
  select(-Plant_ID)

AC_merged_5.5 <- merge(AC_merged_5, leaves_AC_2, by = c("Transect", "Recipient", "Donor", "Gen", "SegPos"), all.x = TRUE) #replicates leaf area & weight for every plant of a transect recipient combo

#Create a column called SLA in the merged dataframe for both individual plant and segement
AC_merged_6 <- AC_merged_5.5 %>%
  mutate(SLA_SEG = leaf_area_mm2_SEG / leaf_dryweight_mg_SEG) %>%
  mutate(SLA = leaf_area_mm2 / leaf_dryweight_mg) %>%
  mutate(LMA_SEG = 1 / SLA_SEG) %>%
  mutate(LMA = 1 / SLA)

```

<br>  


#### d13C data  
[[UPDATED]]

```{r}
AC_d13C_2023 <- read_csv(here::here("data_sheets", "2023_year2_data", "isotope_data", "AC_d13C_2023.csv")) %>% 
  select(1,2,2,3,5,6,9) %>% 
  rename(sample_ID = 1, d13C = 2, totalC_ug = 3, sample_weight_mg = 6)
#Save as UTF-8 file from excel to avoid import error
  
AC_d13C_2023_1.5 <- AC_d13C_2023 %>%
  select(sample_ID, d13C, sample_weight_mg)

AC_d13C_2023_2 <- AC_d13C_2023 %>%
  mutate(sample_ID = substr(sample_ID, 1, nchar(sample_ID) - 2)) %>%
  select(sample_ID, d13C, sample_weight_mg) %>%
  rename(sample_ID_SEG = 1, d13C_SEG = 2, sample_weight_mg_SEG = 3)

AC_merged_6.5 <- AC_merged_6 %>%
  mutate(sample_ID_SEG = paste(Recipient, Gen, str_pad(Transect, width = 3, pad = "0"), SegPos, sep = "-")) %>%
  mutate(sample_ID_SEG = str_replace(sample_ID_SEG, "^([^_]*)_", "\\1-"))

AC_merged_7 <- AC_merged_6.5 %>%
  mutate(sample_ID = paste(Recipient, Gen, str_pad(Transect, width = 3, pad = "0"), SegPos, Plant_ID, sep = "-")) %>%
  mutate(sample_ID = str_replace(sample_ID, "^([^_]*)_", "\\1-"))

#merge a dataframe with isotope data for each plant
AC_merged_7.5 <- merge(AC_merged_7, AC_d13C_2023_1.5, by = c("sample_ID"), all.x = TRUE)

#merge a dataframe with isotope data for each segment
AC_merged_8 <- merge(AC_merged_7.5, AC_d13C_2023_2, by = "sample_ID_SEG", all.x = TRUE)
#Merge by Recipient and Transect only: genotype-wide estimates of d13C since some plants were NF

# Subset to exclude rows where d13C is NA
testdf <- AC_merged_8 %>%
  filter(F_plant == "TRUE") %>% 
filter(!is.na(d13C)) # 1 NF + 416 in the test df = 417 observations

```

<br>  


**Error checks**

```{r}

#Chec for duplicates
AC_dup_check <- AC_merged_8 %>% 
  mutate(duped1 = duplicated(select(AC_merged_8, c(sample_ID))),
         duped2 = duplicated(select(AC_merged_8, c(sample_ID)), fromLast = TRUE),
         duped = case_when(duped1|duped2 == TRUE ~ TRUE,
                           TRUE ~ FALSE)) %>% #highlight all duplicates
  select(-c(duped1, duped2)) %>%
  filter(duped == "TRUE")
# No "TRUE"s found, so no duplicates, no need to remove them

##Check for mismatched (fitness) plants

AC_Echeck1 <- AC_merged_8 %>% 
  filter(F_plant == FALSE & total_fruits != 0) #do the same for seed count, using `>` and `!=`

AC_Echeck2 <- AC_merged_8 %>% 
  filter(F_plant == TRUE) %>% 
  filter(ALL_closed_fruit == 0 & total_seeds > 0)

##NA/NaN/Info values for mean seeds/est seed prod
AC_Echeck3 <- AC_merged_8 %>% 
  filter(F_plant == TRUE & is.na(as.numeric(avg_seeds_per_fruit)))

AC_Echeck4 <- AC_merged_8 %>% 
  filter(F_plant == TRUE & !is.finite(avg_seeds_per_fruit))


## Check for NA seed counts for fitness plants (when fruit production > 0)
#Will be empty when seed_ct NAs are replaced with 0
AC_Echeck5 <- AC_merged_8 %>% 
  filter(F_plant == TRUE & is.na(total_seeds))
##NAs present here before setting seed_ct NAs to 0 were from fitness plants that had 0 closed fruit


#Seed mass
AC_Echeck6 <- AC_merged_8 %>% 
  filter(F_plant == FALSE & !is.na(mean_seed_mass_mg))


```

<br>  


### Trait calculations 

Some trait calculations performed in `Fitness_and_EDA.Rmd.  

Fruits were only collected from fitness plants, so only survival to flowering is calcualted for all plants (fitness and non-fitness).

UPDATE with latest merged df...

```{r}

#Survival measures
#Added var for closed fruit produced
#renamed "survival" to seed production
#
AC_merged_9 <- AC_merged_8 %>% 
  mutate(surv_to_flower = case_when(!is.na(FFD) ~ 1, #plants that survived to flower
                                  is.na(FFD) ~ 0),
         surv_to_fruitprod = case_when(F_plant == TRUE & total_fruits > 0 ~ 1, #plants that survived to produce fruit
                                       F_plant == TRUE & total_fruits == 0 ~ 0),
         prod_closed_fruits = case_when(F_plant == TRUE & ALL_closed_fruit > 0 ~ 1,
                                            F_plant == TRUE & ALL_closed_fruit == 0 ~ 0),
         prod_filled_seeds = case_when(F_plant == TRUE & ALL_filled_seeds > 0 ~ 1, #plants produced filled seeds
                                      F_plant == TRUE & ALL_filled_seeds == 0 ~ 0)) #else = 0

#NOTE
#Adjusting fecundity/fitness:
#The issue of how to handle 0s vs missing data in order to accurately calculate fitness has led us to apply the following corrections:
# Estimated fecundity (seeds per fruit x total fruits) should be set to *0* when closed fruits >= 1 & filled seeds == 0; when closed fruits == open fruits == 0
#Set fecundity to NA when closed = 0 & open >= 1

#Mean seeds per fruit left as-is (NA when no closed fruits collected)
         

AC_merged_10 <- AC_merged_9 %>% 
  rename(mean_seeds_per_fruit = avg_seeds_per_fruit) %>%
  mutate(across(c(est_fecundity, mean_seeds_per_fruit), ~ as.numeric(na_if(., NaN)))) %>% 
#Converts to NA if any of the values are "NaN"
  mutate(mean_seeds_per_fruit = case_when(
          F_plant == TRUE & ALL_closed_fruit == 0 ~ NA, # NA when 0 closed fruits
          TRUE ~ mean_seeds_per_fruit)) %>%  #else, original value
          mutate(est_fecundity = case_when(
          F_plant == TRUE & ALL_closed_fruit == 0 & ALL_open_fruit == 0 ~ 0,
          F_plant == TRUE & ALL_closed_fruit == 0 & ALL_open_fruit >= 1 ~ NA,
          F_plant == TRUE & !is.na(est_fecundity) ~ est_fecundity)) %>%
  select(-sample_ID)


```

<br>  



### Master sheets (2023)  

Full and summarized data sheets for all data collected in the 2022-2023 field season. 
These are merged with the year 1 data sheets elsewhere.  

```{r}
## 'Long' version with all data - one row per plant ID
#Sort by transect and ID variables, then by Fitness plants and those that flowered...

######
######

AC_merged_full_2023 <- AC_merged_10 %>% 
  filter(F_plant == TRUE)
##NOTES - there are still fitness plants that never flowered... keep as 0s
#fitness plants only

write_csv(AC_merged_full_2023, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "AC_merged_full_2023.csv"), na = "")


######
######


AC_mastersheet_v1_2023 <- AC_merged_10 %>% 
  mutate(Year = 2023) %>% 
  select(Year, Gen, Transect, Sequence, Donor, Recipient, Plant_ID, F_plant, everything()) %>% 
  arrange(Transect, Sequence, Recipient, Plant_ID) %>%
  rename(open_fruits = ALL_open_fruit, closed_fruits = ALL_closed_fruit, filled_seeds = ALL_filled_seeds, msm_all = mean_seed_mass_mg)

write_csv(AC_mastersheet_v1_2023, here::here("data_sheets", "compiled_sheets", "AC_mastersheet_full_2023.csv"))
#Renamed to full - has records of all plants (including non-fitness)



#
## Fitness Plants Only
#

AC_mastersheet_F_mains_2023 <-  AC_mastersheet_v1_2023 %>% 
  filter(F_plant == TRUE) #Only main fitness plants

##Note: remaining NA values for fitness measures (mean seeds per fruit; est seed production) are the results of 0 closed fruit counts. Treat these as 0 fitness or as NAs to remove???


write_csv(AC_mastersheet_F_mains_2023, here::here("data_sheets", "compiled_sheets", "AC_mastersheet_Fitness-mains_2023.csv"))

```



