---
title: "BB_22_Vs_23"
author: "Helen Payne"
date: "2024-07-15"
output:
  pdf_document: default
  html_document: default
---

---
output:
  pdf_document: default
  html_document: default
---

```{r}
# Load libraries
library(tidyverse)
library(lme4)
library(car)
```


```{r}
#read in the data
BB_22_23_fit <- read_csv(here::here("data_sheets", "compiled_sheets", "BB_22_23_fit.csv"))

# Create a new column "Genotype" by merging "Donor" and "Recipient" with a hyphen in between
BB_22_23_fit$Genotype <- paste(BB_22_23_fit$Donor, BB_22_23_fit$Recipient, sep = "-")

BB_22_23_full <- read_csv(here::here("data_sheets", "compiled_sheets", "BB_22_23_full.csv"))
```

```{r}

# Calculate the mean surv_to_flower for each group in AC_22_23_full and add it as a new column
BB_22_23_full <- BB_22_23_full %>%
  group_by(Year, Gen, Location, Sequence, Donor, Recipient, SegPos) %>%
  mutate(prop_surv_to_flower = mean(surv_to_flower, na.rm = TRUE)) %>%
  ungroup()  # Remove grouping

BB_prop_sample <- BB_22_23_full %>%
  select(c(Year, Gen, Location, Sequence, Donor, Recipient, SegPos, prop_surv_to_flower))  %>%
    distinct()

BB_22_23_fit <- BB_22_23_fit %>%
  left_join(BB_prop_sample %>% select(Year, Gen, Location, Sequence, Donor, Recipient, SegPos, prop_surv_to_flower),
            by = c("Year", "Gen", "Location", "Sequence", "Donor", "Recipient", "SegPos"))

# Create the new column
BB_22_23_fit <- BB_22_23_fit %>%
  mutate(est_fitness = prop_surv_to_flower * est_fecundity)

```


```{r}
#Subset the data for eBBh year and eBBh generation
BB_22_fit <- BB_22_23_fit %>%
  filter(Year == 2022)

BB_23_fit_G1 <- BB_22_23_fit %>%
  filter(Year == 2023) %>%
  filter(Gen == "G1")

BB_23_fit_G2 <- BB_22_23_fit %>%
  filter(Year == 2023) %>%
  filter(Gen == "G2")

```


Now compare Generation 1 from BB in both 2022 and 2023, but first we need to determine which genotypes are the same in 2022 and 2023

```{r}
# Create a dataframe with common "Genotype" values for 2022
BB_common_22_23_G1_22 <- semi_join(BB_22_fit, BB_23_fit_G1, by = "Genotype")

# Create a dataframe with common "Genotype" values for 2022
BB_common_22_23_G1_23 <- semi_join(BB_23_fit_G1, BB_22_fit, by = "Genotype")

# Merge these two dataframes to have one finalized dataset to compare G1 2022 and G1 2023
BB_common_22_23_G1 <- full_join(BB_common_22_23_G1_22, BB_common_22_23_G1_23, by = intersect(names(BB_common_22_23_G1_22), names(BB_common_22_23_G1_23)))
```

**Determine if 2022 and 2023 trait means are significantly different in Generation 1**
Any difference would be due to environmental effects (because these are the same gneotypes)

#MATERNAL#2022v2023#
```{r}
# Define the traits to analyze
traits <- c("corolla_diam_mm", "skel_dryweight_mg_SEG", "SLA_SEG", "d13C_SEG", "msm_all", "LMA_SEG", "fl_duration", "mean_seeds_per_fruit", "est_fecundity", "est_fitness")

# Loop over eBBh trait
for (trait in traits) {
  
  # Calculate the mean trait value for eBBh maternal (Recipient) family for 2022
  maternal_means_22 <- BB_common_22_23_G1_22 %>%
    group_by(Recipient, Donor) %>%
    summarise(mean_trait = mean(get(trait), na.rm = TRUE))
  
  
  # Calculate the mean trait value for eBBh maternal (Recipient) family for 2023 G1
  maternal_means_23_G1 <- BB_common_22_23_G1_23 %>%
    group_by(Recipient, Donor) %>%
    summarise(mean_trait = mean(get(trait), na.rm = TRUE))
  
  
  # Add year information
  maternal_means_22 <- maternal_means_22 %>%
    mutate(Year = "2022")
  
  maternal_means_23_G1 <- maternal_means_23_G1 %>%
    mutate(Year = "2023_G1")
  
  # Combine the dataframes
  mat_combined_data_year <- bind_rows(maternal_means_22, maternal_means_23_G1)
  
  # Create the histogram
  ggplot(mat_combined_data_year, aes(x = mean_trait, fill = Year)) +
    geom_histogram(position = "identity", alpha = 0.6, binwidth = 0.2) +
    labs(title = paste("Comparison of Recipient Mean", trait),
         x = paste("Mean", trait),
         y = "Frequency") +
    scale_fill_manual(values = c("2022" = "blue", "2023_G1" = "red")) +
    theme_minimal()
  
  # Split the data by year
  trait_2022 <- na.omit(maternal_means_22$mean_trait)
  trait_2023 <- na.omit(maternal_means_23_G1$mean_trait)
  
  # Check normality with Shapiro-Wilk test
  shapiro_test_2022 <- shapiro.test(trait_2022)
  shapiro_test_2023 <- shapiro.test(trait_2023)
  
  # Print Shapiro-Wilk test results
  print(paste("Shapiro-Wilk test for 2022", trait, ":"))
  print(shapiro_test_2022)
  
  print(paste("Shapiro-Wilk test for 2023", trait, ":"))
  print(shapiro_test_2023)
  
  # Perform Levene's test for homogeneity of variances
  levene_test <- leveneTest(mean_trait ~ Year, data = mat_combined_data_year)
  print(paste("Levene's test for", trait, ":"))
  print(levene_test)
  
  # Perform Mann-Whitney U test (Wilcoxon rank-sum test)
  mann_whitney_test <- wilcox.test(mean_trait ~ Year, data = mat_combined_data_year)
  print(paste("Mann-Whitney U test for", trait, ":"))
  print(mann_whitney_test)
  
  # Optionally perform t-test if data is normally distributed
 # if (shapiro_test_2022$p.value > 0.05 && shapiro_test_2023$p.value > 0.05 && levene_test$`Pr(>F)`[1] > 0.05) {
    t_test_result_year <- t.test(mean_trait ~ Year, data = mat_combined_data_year)
    print(paste("t-test for", trait, ":"))
    print(t_test_result_year)
  #} else {
  #  print(paste("Skipping t-test for", trait, "due to non-normal distribution or unequal variances"))
  }

```

#PATERNAL#2022v2023#
```{r}
# Define the traits to analyze
traits <- c("corolla_diam_mm", "skel_dryweight_mg_SEG", "SLA_SEG", "d13C_SEG", "msm_all", "LMA_SEG", "fl_duration", "mean_seeds_per_fruit", "est_fecundity")

# Loop over eBBh trait
for (trait in traits) {
  
  # Calculate the mean trait value for eBBh paternal (Donor) family for 2022
  paternal_means_22 <- BB_common_22_23_G1_22 %>%
    group_by(Donor) %>%
    summarise(mean_trait = mean(get(trait), na.rm = TRUE))
  
  # Calculate the mean trait value for eBBh paternal (Donor) family for 2023 G1
  paternal_means_23_G1 <- BB_common_22_23_G1_23 %>%
    group_by(Donor) %>%
    summarise(mean_trait = mean(get(trait), na.rm = TRUE))
  
  # Add year information
  paternal_means_22 <- paternal_means_22 %>%
    mutate(Year = "2022")
  
  paternal_means_23_G1 <- paternal_means_23_G1 %>%
    mutate(Year = "2023_G1")
  
  # Combine the dataframes
  pat_combined_data_year <- bind_rows(paternal_means_22, paternal_means_23_G1)
  
  # Create the histogram
  ggplot(pat_combined_data_year, aes(x = mean_trait, fill = Year)) +
    geom_histogram(position = "identity", alpha = 0.6, binwidth = 0.2) +
    labs(title = paste("Comparison of Recipient Mean", trait),
         x = paste("Mean", trait),
         y = "Frequency") +
    scale_fill_manual(values = c("2022" = "blue", "2023_G1" = "red")) +
    theme_minimal()
  
  # Split the data by year
  trait_2022 <- na.omit(paternal_means_22$mean_trait)
  trait_2023 <- na.omit(paternal_means_23_G1$mean_trait)
  
  # Check normality with Shapiro-Wilk test
  shapiro_test_2022 <- shapiro.test(trait_2022)
  shapiro_test_2023 <- shapiro.test(trait_2023)
  
  # Print Shapiro-Wilk test results
  print(paste("Shapiro-Wilk test for 2022", trait, ":"))
  print(shapiro_test_2022)
  
  print(paste("Shapiro-Wilk test for 2023", trait, ":"))
  print(shapiro_test_2023)
  
  # Perform Levene's test for homogeneity of variances
  levene_test <- leveneTest(mean_trait ~ Year, data = pat_combined_data_year)
  print(paste("Levene's test for", trait, ":"))
  print(levene_test)
  
  # Perform Mann-Whitney U test (Wilcoxon rank-sum test)
  mann_whitney_test <- wilcox.test(mean_trait ~ Year, data = pat_combined_data_year)
  print(paste("Mann-Whitney U test for", trait, ":"))
  print(mann_whitney_test)
  
  # Optionally perform t-test if data is normally distributed
 # if (shapiro_test_2022$p.value > 0.05 && shapiro_test_2023$p.value > 0.05 && levene_test$`Pr(>F)`[1] > 0.05) {
    t_test_result_year <- t.test(mean_trait ~ Year, data = pat_combined_data_year)
    print(paste("t-test for", trait, ":"))
    print(t_test_result_year)
  #} else {
  #  print(paste("Skipping t-test for", trait, "due to non-normal distribution or unequal variances"))
  }
 # }

```

#CALCULATING PATERNAL FITNESS ESTIMATES
```{r}
# Define the datasets and the years they correspond to
datasets <- list(
  "2022" = BB_common_22_23_G1_22,
  "2023_G1" = BB_common_22_23_G1_23
)

# Initialize an empty list to store results
paternal_means_list <- list()

# Loop over each dataset and perform the calculations
for (year in names(datasets)) {
  # Calculate the mean trait value for each maternal (Recipient) family
  maternal_means <- datasets[[year]] %>%
    group_by(Recipient, Donor) %>%
    summarise(mat_mean_est_fitness = mean(est_fitness, na.rm = TRUE))
  
  # Calculate the mean est_fitness for each Donor
  paternal_means <- maternal_means %>%
    group_by(Donor) %>%
    summarise(pat_mean_est_fitness = mean(mat_mean_est_fitness, na.rm = TRUE)) %>%
    mutate(Year = year)
  
  # Store the result in the list
  paternal_means_list[[year]] <- paternal_means
}

# Combine the dataframes for different years
pat_combined_data_year <- bind_rows(paternal_means_list)

# Split the data by year
trait_2022 <- na.omit(paternal_means_list[["2022"]]$pat_mean_est_fitness)
trait_2023 <- na.omit(paternal_means_list[["2023_G1"]]$pat_mean_est_fitness)

# Check normality with Shapiro-Wilk test
shapiro_test_2022 <- shapiro.test(trait_2022)
shapiro_test_2023 <- shapiro.test(trait_2023)

# Print Shapiro-Wilk test results
print("Shapiro-Wilk test for 2022:")
print(shapiro_test_2022)

print("Shapiro-Wilk test for 2023:")
print(shapiro_test_2023)

# Perform Levene's test for homogeneity of variances
levene_test <- leveneTest(pat_mean_est_fitness ~ Year, data = pat_combined_data_year)
print("Levene's test:")
print(levene_test)

# Perform Mann-Whitney U test (Wilcoxon rank-sum test)
mann_whitney_test <- wilcox.test(pat_mean_est_fitness ~ Year, data = pat_combined_data_year)
print("Mann-Whitney U test:")
print(mann_whitney_test)

# Optionally perform t-test if data is normally distributed
# if (shapiro_test_2022$p.value > 0.05 && shapiro_test_2023$p.value > 0.05 && levene_test$`Pr(>F)`[1] > 0.05) {
t_test_result_year <- t.test(pat_mean_est_fitness ~ Year, data = pat_combined_data_year)
print("t-test:")
print(t_test_result_year)
# } else {
#   print("Skipping t-test due to non-normal distribution or unequal variances")
# }
```

#PHENOTYPE#2022v2023
```{r}
# Define the traits to analyze
traits <- c("corolla_diam_mm", "skel_dryweight_mg_SEG", "SLA_SEG", "d13C_SEG", "msm_all", "LMA_SEG", "fl_duration", "mean_seeds_per_fruit", "est_fecundity", "est_fitness")

# Loop over eBBh trait
for (trait in traits) {
  
  # Convert 'Year' to factor if it's not already
  if (!is.factor(BB_common_22_23_G1$Year)) {
    BB_common_22_23_G1$Year <- as.factor(BB_common_22_23_G1$Year)
  }
  
  # Create the histogram
  ggplot(BB_common_22_23_G1, aes_string(x = trait, fill = "Year")) +
    geom_histogram(position = "identity", alpha = 0.6, binwidth = 0.2) +
    labs(title = paste("Comparison of", trait),
         x = trait,
         y = "Frequency") +
    scale_fill_manual(values = c("2022" = "blue", "2023" = "red")) +
    theme_minimal()
  
  # Split the data by year
  trait_2022 <- na.omit(BB_common_22_23_G1_22[[trait]])
  trait_2023 <- na.omit(BB_common_22_23_G1_23[[trait]])
  
  # Check normality with Shapiro-Wilk test
  shapiro_test_2022 <- shapiro.test(trait_2022)
  shapiro_test_2023 <- shapiro.test(trait_2023)
  
  # Print Shapiro-Wilk test results
  print(paste("Shapiro-Wilk test for 2022", trait, ":"))
  print(shapiro_test_2022)
  
  print(paste("Shapiro-Wilk test for 2023", trait, ":"))
  print(shapiro_test_2023)
  
  # Perform Levene's test for homogeneity of variances
  levene_test <- leveneTest(as.formula(paste(trait, "~ as.factor(Year)")), data = BB_common_22_23_G1)
  print(paste("Levene's test for", trait, ":"))
  print(levene_test)
  
  # Perform Mann-Whitney U test (Wilcoxon rank-sum test)
  mann_whitney_test <- wilcox.test(as.formula(paste(trait, "~ Year")), data = BB_common_22_23_G1)
  print(paste("Mann-Whitney U test for", trait, ":"))
  print(mann_whitney_test)
  
  # Optionally perform t-test if data is normally distributed and variances are equal
  #if (shapiro_test_2022$p.value > 0.05 && shapiro_test_2023$p.value > 0.05 && levene_test$`Pr(>F)`[1] > 0.05) {
    t_test_result_year <- t.test(as.formula(paste(trait, "~ Year")), data = BB_common_22_23_G1)
    print(paste("t-test for", trait, ":"))
    print(t_test_result_year)
  } #else {
    #print(paste("Skipping t-test for", trait, "due to non-normal distribution or unequal variances"))
  #}
#}
```



**Determine if G1 and G2 trait means are significantly different in 2023**
Any difference would be due to either the effect of greenhouse produced seeds or a generation of selection

#MATERNAL#G1vG2#
```{r}
# Define the traits to analyze
traits <- c("corolla_diam_mm_SEG", "skel_dryweight_mg_SEG", "SLA_SEG", "d13C_SEG", "msm_all", "LMA_SEG", "fl_duration", "mean_seeds_per_fruit", "est_fecundity", "est_fitness")

# Loop over eBBh trait
for (trait in traits) {
  
  # Create a dataframe with common "Genotype" values for 2023
  BB_common_23_G1 <- semi_join(BB_23_fit_G1, BB_23_fit_G2, by = "Genotype")
  BB_common_23_G2 <- semi_join(BB_23_fit_G2, BB_23_fit_G1, by = "Genotype")
  
  # Merge these two dataframes to have one finalized dataset to compare G1 2023 and G2 2023
  BB_common_23_G1_G2 <- full_join(BB_common_23_G1, BB_common_23_G2, by = intersect(names(BB_common_23_G1), names(BB_common_23_G2)))

  # Calculate the mean trait value for eBBh maternal (Recipient) family for G1, 2023
  maternal_means_23_G1 <- BB_common_23_G1 %>%
    group_by(Recipient, Donor) %>%
    summarise(mean_trait = mean(get(trait), na.rm = TRUE)) %>%
    mutate(Gen = "1")
  
  # Calculate the mean trait value for eBBh maternal (Recipient) family for G2, 2023
  maternal_means_23_G2 <- BB_common_23_G2 %>%
    group_by(Recipient, Donor) %>%
    summarise(mean_trait = mean(get(trait), na.rm = TRUE)) %>%
    mutate(Gen = "2")
  
  # Combine the dataframes
  mat_combined_data_gen <- bind_rows(maternal_means_23_G1, maternal_means_23_G2)

  # Create the histogram
  ggplot(mat_combined_data_gen, aes(x = mean_trait, fill = Gen)) +
    geom_histogram(position = "identity", alpha = 0.6, binwidth = 0.2) +
    labs(title = paste("Comparison of Recipient Mean", trait, ": G1 vs G2"),
         x = paste("Mean", trait),
         y = "Frequency") +
    scale_fill_manual(values = c("1" = "blue", "2" = "red")) +
    theme_minimal()
  
  # Split the data by generation
  trait_G1 <- na.omit(maternal_means_23_G1$mean_trait)
  trait_G2 <- na.omit(maternal_means_23_G2$mean_trait)
  
  # Check normality with Shapiro-Wilk test
  shapiro_test_G1 <- shapiro.test(trait_G1)
  shapiro_test_G2 <- shapiro.test(trait_G2)
  
  # Print Shapiro-Wilk test results
  print(shapiro_test_G1)
  print(shapiro_test_G2)
  
  # Perform Levene's test for homogeneity of variances
  levene_test <- leveneTest(mean_trait ~ as.factor(Gen), data = mat_combined_data_gen)
  
  # Print Levene's test result
  print(levene_test)
  
  # Perform Mann-Whitney U test (Wilcoxon rank-sum test) instead of t-test
  mann_whitney_test <- wilcox.test(mean_trait ~ Gen, data = mat_combined_data_gen)
  
  # Print Mann-Whitney test result
  print(mann_whitney_test)
  
  # Perform t-test
  t_test_result_gen <- t.test(mean_trait ~ Gen, data = mat_combined_data_gen)
  
  # Print t-test result
  print(t_test_result_gen)

}

```

#PATERNAL#G1vG2#
```{r}
# Define the traits to analyze
traits <- c("corolla_diam_mm_SEG", "skel_dryweight_mg_SEG", "SLA_SEG", "d13C_SEG", "msm_all", "LMA_SEG", "fl_duration", "mean_seeds_per_fruit", "est_fecundity")

# Loop over eBBh trait
for (trait in traits) {
  
  # Create a dataframe with common "Genotype" values for 2023
  BB_common_23_G1 <- semi_join(BB_23_fit_G1, BB_23_fit_G2, by = "Genotype")
  BB_common_23_G2 <- semi_join(BB_23_fit_G2, BB_23_fit_G1, by = "Genotype")
  
  # Merge these two dataframes to have one finalized dataset to compare G1 2023 and G2 2023
  BB_common_23_G1_G2 <- full_join(BB_common_23_G1, BB_common_23_G2, by = intersect(names(BB_common_23_G1), names(BB_common_23_G2)))

  # Calculate the mean trait value for eBBh paternal (Donor) family for G1, 2023
  paternal_means_23_G1 <- BB_common_23_G1 %>%
    group_by(Donor) %>%
    summarise(mean_trait = mean(get(trait), na.rm = TRUE), .groups = "drop") %>%
    mutate(Gen = "1")
  
  # Calculate the mean trait value for eBBh paternal (Donor) family for G2, 2023
  paternal_means_23_G2 <- BB_common_23_G2 %>%
    group_by(Donor) %>%
    summarise(mean_trait = mean(get(trait), na.rm = TRUE), .groups = "drop") %>%
    mutate(Gen = "2")
  
  # Combine the dataframes
  pat_combined_data_gen <- bind_rows(paternal_means_23_G1, paternal_means_23_G2)

  # Create the histogram
  ggplot(pat_combined_data_gen, aes(x = mean_trait, fill = Gen)) +
    geom_histogram(position = "identity", alpha = 0.6, binwidth = 0.2) +
    labs(title = paste("Comparison of Donor Mean", trait, ": G1 vs G2"),
         x = paste("Mean", trait),
         y = "Frequency") +
    scale_fill_manual(values = c("1" = "blue", "2" = "red")) +
    theme_minimal()
  
  # Split the data by generation and ensure numeric conversion
  trait_G1 <- na.omit(as.numeric(paternal_means_23_G1$mean_trait))
  trait_G2 <- na.omit(as.numeric(paternal_means_23_G2$mean_trait))
  
  # Check normality with Shapiro-Wilk test if sample size is appropriate
  if (length(trait_G1) >= 3 && length(trait_G1) <= 5000) {
    shapiro_test_G1 <- shapiro.test(trait_G1)
    print(shapiro_test_G1)
  } else {
    print(paste("Sample size for trait_G1 is not suitable for Shapiro-Wilk test:", length(trait_G1)))
  }
  
  if (length(trait_G2) >= 3 && length(trait_G2) <= 5000) {
    shapiro_test_G2 <- shapiro.test(trait_G2)
    print(shapiro_test_G2)
  } else {
    print(paste("Sample size for trait_G2 is not suitable for Shapiro-Wilk test:", length(trait_G2)))
  }
  
  # Perform Levene's test for homogeneity of variances
  levene_test <- leveneTest(mean_trait ~ as.factor(Gen), data = pat_combined_data_gen)
  print(levene_test)
  
  # Perform Mann-Whitney U test (Wilcoxon rank-sum test)
  mann_whitney_test <- wilcox.test(mean_trait ~ Gen, data = pat_combined_data_gen)
  print(mann_whitney_test)
  
  # Perform t-test
  t_test_result_gen <- t.test(mean_trait ~ Gen, data = pat_combined_data_gen)
  print(t_test_result_gen)
}

```

#CALCULATING PATERNAL FITNESS ESTIMATES
```{r}
# Define the datasets and the years they correspond to
datasets <- list(
  "2023_G1" = BB_common_23_G1,
  "2023_G2" = BB_common_23_G2
)

# Initialize an empty list to store results
paternal_means_list <- list()

# Loop over each dataset and perform the calculations
for (year in names(datasets)) {
  # Calculate the mean trait value for each maternal (Recipient) family
  maternal_means <- datasets[[year]] %>%
    group_by(Recipient, Donor) %>%
    summarise(mat_mean_est_fitness = mean(est_fitness, na.rm = TRUE))
  
  # Calculate the mean est_fitness for each Donor
  paternal_means <- maternal_means %>%
    group_by(Donor) %>%
    summarise(pat_mean_est_fitness = mean(mat_mean_est_fitness, na.rm = TRUE)) %>%
    mutate(Year = year)
  
  # Store the result in the list
  paternal_means_list[[year]] <- paternal_means
}

# Combine the dataframes for different years
pat_combined_data_year <- bind_rows(paternal_means_list)

# Split the data by year
trait_2023_G1 <- na.omit(paternal_means_list[["2023_G1"]]$pat_mean_est_fitness)
trait_2023_G2 <- na.omit(paternal_means_list[["2023_G2"]]$pat_mean_est_fitness)

# Check normality with Shapiro-Wilk test
shapiro_test_2023_G1 <- shapiro.test(trait_2023_G1)
shapiro_test_2023_G2 <- shapiro.test(trait_2023_G2)

# Print Shapiro-Wilk test results
print("Shapiro-Wilk test for 2023 G1:")
print(shapiro_test_2023_G1)

print("Shapiro-Wilk test for 2023 G2:")
print(shapiro_test_2023_G2)

# Perform Levene's test for homogeneity of variances
levene_test <- leveneTest(pat_mean_est_fitness ~ Year, data = pat_combined_data_year)
print("Levene's test:")
print(levene_test)

# Perform Mann-Whitney U test (Wilcoxon rank-sum test)
mann_whitney_test <- wilcox.test(pat_mean_est_fitness ~ Year, data = pat_combined_data_year)
print("Mann-Whitney U test:")
print(mann_whitney_test)

# Optionally perform t-test if data is normally distributed
# if (shapiro_test_2022$p.value > 0.05 && shapiro_test_2023$p.value > 0.05 && levene_test$`Pr(>F)`[1] > 0.05) {
t_test_result_year <- t.test(pat_mean_est_fitness ~ Year, data = pat_combined_data_year)
print("t-test:")
print(t_test_result_year)
# } else {
#   print("Skipping t-test due to non-normal distribution or unequal variances")
# }
```

#PHENOTYPE#G1vG2
```{r}
# Define the traits to analyze
traits <- c("corolla_diam_mm_SEG", "skel_dryweight_mg_SEG", "SLA_SEG", "d13C_SEG", "msm_all", "LMA_SEG", "fl_duration", "mean_seeds_per_fruit", "est_fecundity", "est_fitness")

# Loop over eBBh trait
for (trait in traits) {
  
  # Create a dataframe with common "Genotype" values for 2023
  BB_common_23_G1 <- semi_join(BB_23_fit_G1, BB_23_fit_G2, by = "Genotype")
  BB_common_23_G2 <- semi_join(BB_23_fit_G2, BB_23_fit_G1, by = "Genotype")
  
  # Merge these two dataframes to have one finalized dataset to compare G1 2023 and G2 2023
  BB_common_23_G1_G2 <- full_join(BB_common_23_G1, BB_common_23_G2, by = intersect(names(BB_common_23_G1), names(BB_common_23_G2)))
  
  # Convert 'Gen' to factor if it's not already
  if (!is.factor(BB_common_23_G1_G2$Gen)) {
    BB_common_23_G1_G2$Gen <- as.factor(BB_common_23_G1_G2$Gen)
  }
  
  # Create the histogram
  ggplot(BB_common_23_G1_G2, aes_string(x = trait, fill = "Gen")) +
    geom_histogram(position = "identity", alpha = 0.6, binwidth = 0.2) +
    labs(title = paste("Comparison of", trait),
         x = trait,
         y = "Frequency") +
    scale_fill_manual(values = c("1" = "blue", "2" = "red")) +
    theme_minimal()
  
  # Split the data by generation and ensure numeric conversion
  trait_G1 <- na.omit(as.numeric(BB_common_23_G1[[trait]]))
  trait_G2 <- na.omit(as.numeric(BB_common_23_G2[[trait]]))
  
  # Check normality with Shapiro-Wilk test if sample size is appropriate
  if (length(trait_G1) >= 3 && length(trait_G1) <= 5000) {
    shapiro_test_G1 <- shapiro.test(trait_G1)
    print(paste("Shapiro-Wilk test for G1", trait, ":"))
    print(shapiro_test_G1)
  } else {
    print(paste("Sample size for G1", trait, "is not suitable for Shapiro-Wilk test:", length(trait_G1)))
  }
  
  if (length(trait_G2) >= 3 && length(trait_G2) <= 5000) {
    shapiro_test_G2 <- shapiro.test(trait_G2)
    print(paste("Shapiro-Wilk test for G2", trait, ":"))
    print(shapiro_test_G2)
  } else {
    print(paste("Sample size for G2", trait, "is not suitable for Shapiro-Wilk test:", length(trait_G2)))
  }
  
  # Perform Levene's test for homogeneity of variances
  levene_test <- leveneTest(as.formula(paste(trait, "~ as.factor(Gen)")), data = BB_common_23_G1_G2)
  print(paste("Levene's test for", trait, ":"))
  print(levene_test)
  
  # Perform Mann-Whitney U test (Wilcoxon rank-sum test)
  mann_whitney_test <- wilcox.test(as.formula(paste(trait, "~ Gen")), data = BB_common_23_G1_G2)
  print(paste("Mann-Whitney U test for", trait, ":"))
  print(mann_whitney_test)
  
  # Perform t-test
  t_test_result_gen <- t.test(as.formula(paste(trait, "~ Gen")), data = BB_common_23_G1_G2)
  print(paste("t-test for", trait, ":"))
  print(t_test_result_gen)
}

```


