---
title: "Data_Snooping_AC"
author: "Helen Payne"
date: "2024-06-27"
output:
  pdf_document: default
  html_document: default
---

```{r}
# Load necessary library
library(tidyverse)
library(lme4)

```

**DATA SNOOPING: looking into survival**
Next we wanted to add in survival into AC_22_23_fit, but in order to do this, we need to get the data from AC_22_23_full
```{r}
#read in both datasets
AC_22_23_full <- read_csv(here::here("data_sheets", "compiled_sheets", "AC_22_23_full.csv"))

AC_22_23_fit <- read_csv(here::here("data_sheets", "compiled_sheets", "AC_22_23_fit.csv"))


# Calculate the mean surv_to_flower for each group in AC_22_23_full and add it as a new column
AC_22_23_full <- AC_22_23_full %>%
  group_by(Year, Gen, Transect, Sequence, Donor, Recipient, SegPos) %>%
  mutate(prop_surv_to_flower = mean(surv_to_flower, na.rm = TRUE)) %>%
  ungroup()  # Remove grouping

AC_prop_sample <- AC_22_23_full %>%
  select(c(Year, Gen, Transect, Sequence, Donor, Recipient, SegPos, prop_surv_to_flower))  %>%
    distinct()

AC_22_23_fit <- AC_22_23_fit %>%
  left_join(AC_prop_sample %>% select(Year, Gen, Transect, Sequence, Donor, Recipient, SegPos, prop_surv_to_flower),
            by = c("Year", "Gen", "Transect", "Sequence", "Donor", "Recipient", "SegPos"))
```

SURVIVAL:
```{r}
# Create the data snooping scatterplot
ggplot(data = AC_22_23_fit, aes(x = corolla_diam_mm_SEG, y = prop_surv_to_flower)) +
  geom_point() +
  labs(
    title = "Scatter Plot of Corolla Diameter vs. Proportion Surviving to Flower",
    x = "Corolla Diameter (mm)",
    y = "Proportion Surviving to Flower"
  ) +
  theme_minimal()


# Create the data snooping scatterplot
ggplot(data = AC_22_23_fit, aes(x = skel_dryweight_mg_SEG, y = prop_surv_to_flower)) +
  geom_point() +
  labs(
    title = "Scatter Plot of Corolla Diameter vs. Proportion Surviving to Flower",
    x = "Skeleton Weight (mg)",
    y = "Proportion Surviving to Flower"
  ) +
  theme_minimal()

# Create the data snooping scatterplot
ggplot(data = AC_22_23_fit, aes(x = mean_seeds_per_fruit, y = prop_surv_to_flower)) +
  geom_point() +
  labs(
    title = "Scatter Plot of Corolla Diameter vs. Proportion Surviving to Flower",
    x = "Mean Seeds Per Fruit",
    y = "Proportion Surviving to Flower"
  ) +
  theme_minimal()

# Create the data snooping scatterplot
ggplot(data = AC_22_23_fit, aes(x = SLA_SEG, y = prop_surv_to_flower)) +
  geom_point() +
  labs(
    title = "Scatter Plot of Corolla Diameter vs. Proportion Surviving to Flower",
    x = "SLA",
    y = "Proportion Surviving to Flower"
  ) +
  theme_minimal()

# Create the data snooping scatterplot
ggplot(data = AC_22_23_fit, aes(x = LMA_SEG, y = prop_surv_to_flower)) +
  geom_point() +
  labs(
    title = "Scatter Plot of Corolla Diameter vs. Proportion Surviving to Flower",
    x = "LMA",
    y = "Proportion Surviving to Flower"
  ) +
  theme_minimal()

# Create the data snooping scatterplot
ggplot(data = AC_22_23_fit, aes(x = d13C_SEG, y = prop_surv_to_flower)) +
  geom_point() +
  labs(
    title = "Scatter Plot of Corolla Diameter vs. Proportion Surviving to Flower",
    x = "d13C",
    y = "Proportion Surviving to Flower"
  ) +
  theme_minimal()
```


ESTIMATED LIFETIME FECUNDITY
```{r}
# Create the data snooping scatterplot
ggplot(data = AC_22_23_fit, aes(x = corolla_diam_mm_SEG, y = est_fecundity)) +
  geom_point() +
  labs(
    title = "Corolla Diameter vs. Proportion Surviving to Flower",
    x = "Corolla Diameter (mm)",
    y = "Estimated Fecundity"
  ) +
  theme_minimal()


# Create the data snooping scatterplot
ggplot(data = AC_22_23_fit, aes(x = skel_dryweight_mg_SEG, y = est_fecundity)) +
  geom_point() +
  labs(
    title = "Skeleton Weight vs. Proportion Surviving to Flower",
    x = "Skeleton Weight (mg)",
    y = "Estimated Fecundity"
  ) +
  theme_minimal()

# Create the data snooping scatterplot
ggplot(data = AC_22_23_fit, aes(x = mean_seeds_per_fruit, y = est_fecundity)) +
  geom_point() +
  labs(
    title = "Mean Seeds Per Fruit vs. Proportion Surviving to Flower",
    x = "Mean Seeds Per Fruit",
    y = "Estimated Fecundity"
  ) +
  theme_minimal()

# Create the data snooping scatterplot
ggplot(data = AC_22_23_fit, aes(x = SLA_SEG, y = est_fecundity)) +
  geom_point() +
  labs(
    title = "SLA vs. Proportion Surviving to Flower",
    x = "SLA",
    y = "Estimated Fecundity"
  ) +
  theme_minimal()

# Create the data snooping scatterplot
ggplot(data = AC_22_23_fit, aes(x = LMA_SEG, y = est_fecundity)) +
  geom_point() +
  labs(
    title = "LMA vs. Proportion Surviving to Flower",
    x = "LMA",
    y = "Estimated Fecundity"
  ) +
  theme_minimal()

# Create the data snooping scatterplot
ggplot(data = AC_22_23_fit, aes(x = d13C_SEG, y = est_fecundity)) +
  geom_point() +
  labs(
    title = "d13C vs. Proportion Surviving to Flower",
    x = "delta Carbon 13",
    y = "Estimated Fecundity"
  ) +
  theme_minimal()
```

OLS LM: Donor, Recipient, Transect
```{r}
#corolla_daim and Skel dryweight model
mod_1 <- lmer(est_fecundity ~ corolla_diam_mm + skel_dryweight_mg + (1|Transect)  + (1 | Donor)  + (1 | Donor: Recipient), data = AC_22_23_fit)

residuals <- resid(mod_1)

# Q-Q plot for normality
qqnorm(residuals) #good


# Histogram for normality
hist(residuals, breaks = 20, main = "Histogram of Residuals", xlab = "Residuals") #normal

AIC(mod_1)
mod_1

# now include SLA
mod_2 <- lmer(est_fecundity ~ corolla_diam_mm_SEG + skel_dryweight_mg_SEG + SLA_SEG + (1|Transect)  + (1 | Donor)  + (1 | Donor: Recipient), data = AC_22_23_fit)

residuals <- resid(mod_2)

# Q-Q plot for normality
qqnorm(residuals) #good


# Histogram for normality
hist(residuals, breaks = 20, main = "Histogram of Residuals", xlab = "Residuals") #normal

AIC(mod_2)
mod_2

# now include d13C
mod_3 <- lmer(est_fecundity ~ corolla_diam_mm_SEG + skel_dryweight_mg_SEG + SLA_SEG + d13C + (1|Transect)  + (1 | Donor) + (1 | Donor: Recipient), data = AC_22_23_fit)

residuals <- resid(mod_3)

# Q-Q plot for normality
qqnorm(residuals) #good


# Histogram for normality
hist(residuals, breaks = 20, main = "Histogram of Residuals", xlab = "Residuals") #normal

AIC(mod_3)
mod_3

# now include SLA
mod_4 <- lmer(est_fecundity ~ corolla_diam_mm_SEG + skel_dryweight_mg_SEG + SLA_SEG + d13C + msm_all + (1|Transect)  + (1 | Donor) + (1 | Donor: Recipient), data = AC_22_23_fit)

residuals <- resid(mod_4)

# Q-Q plot for normality
qqnorm(residuals) #good


# Histogram for normality
hist(residuals, breaks = 20, main = "Histogram of Residuals", xlab = "Residuals") #normal

AIC(mod_4)
BIC(mod_4)
mod_4
```

**Next Step:**
Using the regression coefficients to calculate the selection gradient, then multiplying this by the phenotypic variance to determine the selection differential for each trait
```{r}
# Extract the fixed effects coefficients
mod_coef <- fixef(mod_4)


# Display the coefficients
print(mod_coef)

# Calculate variances of the traits
corolla_var <- var(AC_22_23_fit$corolla_diam_mm_SEG, na.rm = TRUE)
skel_var <- var(AC_22_23_fit$skel_dryweight_mg_SEG, na.rm = TRUE)
SLA_var <- var(AC_22_23_fit$SLA_SEG, na.rm = TRUE)
d13C_var <- var(AC_22_23_fit$d13C_SEG, na.rm = TRUE)
msm_var <- var(AC_22_23_fit$msm_all, na.rm = TRUE)

# Ensure the coefficients exist before calculation
corolla_selection_diff <- ifelse(!is.na(mod_coef["corolla_diam_mm_SEG"]), mod_coef["corolla_diam_mm_SEG"] * corolla_var, NA)
skel_selection_diff <- ifelse(!is.na(mod_coef["skel_dryweight_mg_SEG"]), mod_coef["skel_dryweight_mg_SEG"] * skel_var, NA)
SLA_diff <- ifelse(!is.na(mod_coef["SLA_SEG"]), mod_coef["SLA_SEG"] * SLA_var, NA)
d13C_diff <- ifelse(!is.na(mod_coef["d13C_SEG"]), mod_coef["d13C_SEG"] * d13C_var, NA)
msm_diff <- ifelse(!is.na(mod_coef["msm_all"]), mod_coef["msm_all"] * msm_var, NA)

# Print selection differentials
print(corolla_selection_diff)
print(skel_selection_diff)
print(SLA_diff)
print(d13C_diff)
print(msm_diff)

# Combine into a dataframe and handle NA values
selection_diff_mod4 <- data.frame(
  Trait = c("corolla_diam_mm", "skel_dryweight_mg", "SLA", "d13C", "msm"),
  Selection_Differential = c(corolla_selection_diff, skel_selection_diff, SLA_diff, d13C_diff, msm_diff)
)

# Remove rows with NA values
selection_diff_mod4 <- selection_diff_mod4[complete.cases(selection_diff_mod4), ]

# Print the dataframe
print(selection_diff_mod4)

```


OLS LM: Donor, Transect
```{r}
mod_5 <- lmer(est_fecundity ~ corolla_diam_mm_SEG + skel_dryweight_mg_SEG + SLA_SEG + d13C + msm_all + (1|Transect)  + (1 | Donor) , data = AC_22_23_fit)

residuals <- resid(mod_5)

# Q-Q plot for normality
qqnorm(residuals) #good


# Histogram for normality
hist(residuals, breaks = 20, main = "Histogram of Residuals", xlab = "Residuals") #normal


AIC(mod_5) #very similar to mod 4, but slightly better
BIC(mod_5)
mod_5
```


OLS LM: Transect
```{r}
mod_6 <- lm(est_fecundity ~ corolla_diam_mm_SEG + skel_dryweight_mg_SEG + SLA_SEG + d13C + msm_all + (1|Transect), data = AC_22_23_fit)

residuals <- resid(mod_6)

# Q-Q plot for normality
qqnorm(residuals) #good


# Histogram for normality
hist(residuals, breaks = 20, main = "Histogram of Residuals", xlab = "Residuals") #normal


AIC(mod_6) #very similar to mod 4 and 5
BIC(mod_6)
mod_6
```

OLS LM: Transect and Recipient
```{r}
mod_7 <- lmer(est_fecundity ~ corolla_diam_mm_SEG + skel_dryweight_mg_SEG + SLA_SEG + d13C + msm_all + (1|Transect) + (1|Recipient), data = AC_22_23_fit)

residuals <- resid(mod_7)

# Q-Q plot for normality
qqnorm(residuals) #good


# Histogram for normality
hist(residuals, breaks = 20, main = "Histogram of Residuals", xlab = "Residuals") #normal


AIC(mod_7) #very similar to mod 4 and 5 and 6
BIC(mod_7)
mod_7
```



```{r}
# Extract the fixed effects coefficients
mod_coef <- fixef(mod_5)

# Display the coefficients
print(mod_coef)

# Calculate variances of the traits
corolla_var <- var(AC_22_23_fit$corolla_diam_mm, na.rm = TRUE)
skel_var <- var(AC_22_23_fit$skel_dryweight_mg, na.rm = TRUE)
SLA_var <- var(AC_22_23_fit$SLA_SEG, na.rm = TRUE)
d13C_var <- var(AC_22_23_fit$d13C_SEG, na.rm = TRUE)
msm_var <- var(AC_22_23_fit$msm_all, na.rm = TRUE)

# Ensure the coefficients exist before calculation
corolla_selection_diff <- ifelse(!is.na(mod_coef["corolla_diam_mm_SEG"]), mod_coef["corolla_diam_mm_SEG"] * corolla_var, NA)
skel_selection_diff <- ifelse(!is.na(mod_coef["skel_dryweight_mg_SEG"]), mod_coef["skel_dryweight_mg_SEG"] * skel_var, NA)
SLA_diff <- ifelse(!is.na(mod_coef["SLA_SEG"]), mod_coef["SLA_SEG"] * SLA_var, NA)
d13C_diff <- ifelse(!is.na(mod_coef["d13C_SEG"]), mod_coef["d13C_SEG"] * d13C_var, NA)
msm_diff <- ifelse(!is.na(mod_coef["msm_all"]), mod_coef["msm_all"] * msm_var, NA)

# Print selection differentials
print(corolla_selection_diff)
print(skel_selection_diff)
print(SLA_diff)
print(d13C_diff)
print(msm_diff)

# Combine into a dataframe
selection_diff_mod5 <- data.frame(
  traits = c("corolla_diam_mm", "skel_dryweight_mg", "SLA", "d13C", "msm"),
  Selection_Differential = c(corolla_selection_diff, skel_selection_diff, SLA_diff, d13C_diff, msm_diff)
)

# Print the dataframe
print(selection_diff_mod5)
```


```{r}
variance_components_AC <- read_csv(here::here("data_sheets", "compiled_sheets", "AC_Va_h2_R.csv")) %>%
    merge(selection_diff_mod4, by = traits)

```

