---
title: "compile_datasheets_BO_2023"
author: "Helen Payne and Devin Gamble"
date: "2024-23-05"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message = FALSE, warning = FALSE}
#Load Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(reshape2)
library(stringr)
library(dplyr)

```



### Load and Clean Data   


**Load Fitness Plant Data**  

- Add in Donor information
```{r}
BO_Fitness <- read_csv(here::here("data_sheets", "2023_year2_data", "fitness_plants", "BO_fitnessplants_23.csv")) %>% 
  slice(-620)



#Donor info loaded from 2021-22 data
BO_donors <- read_csv(here::here("data_sheets", "tray_positions_2021_year1","BO_tray_rand.csv"), skip = 1)

# Use dplyr::select to avoid conflicts
BO_donors <- BO_donors %>%
  dplyr::select(Donor_ID, Recipient_ID) %>%
  rename(Donor = Donor_ID, Recipient = Recipient_ID)



#Fix Recipient and Donor names
BO_donors2 <- BO_donors %>% 
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% 
  mutate(Recipient = as.character(str_c("BO_", Recipient_num))) %>% 
  mutate(Donor_num = sprintf("%03d", (parse_number(Donor)))) %>% 
  mutate(Donor = as.character(str_c("BO_", Donor_num))) %>% 
  dplyr::select(-Recipient_num, -Donor_num)

#Multiple matches, merge only unique pairs for Donor
BO_Fitness_2 <- merge(BO_Fitness, unique(BO_donors2[c("Donor", "Recipient")]), by = c("Recipient"), all.x = TRUE) %>% 
  mutate(SegPos = paste(Segment, position, sep = "")) %>% 
  dplyr::select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Fit_plant) %>% 
  arrange(Sequence, SegPos)

#Save as V2 Fitness Plant Sheet
write_csv(BO_Fitness_2, here::here("data_sheets", "2023_year2_data", "fitness_plants", "BO_fitnessplants_23_V2.csv"), na = "")


```

<br>  


### Trait Data  

Trait data gathered and processed from the 2022-2023 field season is loaded and organized below.  


#### Flower Check & Phenology  
-- Organize flowering data
-- Add in Sow & Plant dates

#########
Crucial Step!!  
######### 
In the csv file, dplyr::select all data with dates (visit dates, leaf sampling, and flower photos). Right click --> format cells. dplyr::select the date category --> YYYY-MM-DD. Save the csv and close it. This will standardize all dates to the same format and make working with them in R (and lubridate) much easier!  

```{r}

BO_flr_df <- read_csv(file = here::here( "data_sheets", "2023_year2_data", "flowering_checks", "BO_flowercheck_data_2023-02-24 - BO_flowercheck_data_v1.csv"), na = c("", "NA")) %>% 
  rename(F_Num_02 = 19)


BO_flr2 <- BO_flr_df %>% 
  mutate(SegPos = paste(Segment, side, sep = "")) %>% 
  dplyr::select(Sequence, Transect, SegPos, everything(), -c(Segment, side, FD_06, FD_07)) %>% 
  slice(-c(698:700, 779:780, 3096:3102)) %>% #Remove individuals not planted and NA rows
  arrange(Sequence, Transect, SegPos)
  

#Check for mismatched fitness plants
BO_flr_test <- BO_flr2 %>% 
  filter(!is.na(Fit_plant)) %>% 
  mutate(FitMatch = case_when(Fit_plant == Plant_ID ~ TRUE,
                              Fit_plant != Plant_ID ~ FALSE))


## Fix Dates!

#Check obs first using unique(BO_flr3$FD_01) and make corrections

BO_flr3 <- BO_flr2 %>%
  mutate_at(.vars = vars(FD_01:FLR_P), ~ as_date(.))

```


Add sowing/planting date  
```{r}
#Sow Date - first date (of two or one) on which seeds were sown in the lab
#Plant Date - median date on which seed plugs were planted in the field

BO_flr4 <- BO_flr3 %>% 
  mutate(Sow_Date = as_date("2022-10-31"),
         Plant_Date = as_date("2022-11-03"))

```


Merge in Fitness plant Data  
- Donor
```{r}
#Check to make sure fitness plants line up
BO_flr5 <- merge(BO_flr4, BO_Fitness_2, by = c("Sequence", "Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE) %>%
  arrange(Sequence, Transect, SegPos, Plant_ID) %>% 
  mutate(F_plant = case_when(!is.na(Fit_plant.x) ~ TRUE,
                               is.na(Fit_plant.x) ~ FALSE)) %>% 
  dplyr::select(-c(Fit_plant.x, Fit_plant.y)) #Drop extra fitplant row after checking that they all match - retain multiple Donors per segment
  
  
#####  
#Check to make sure fitness plants line up
#BO_flr_test <- BO_flr5 %>% 
#  filter(!is.na(Fit_plant.x)) %>%
#   mutate(FMatch = case_when(Fit_plant.x == Fit_plant.y ~ TRUE,
#                           Fit_plant.x != Fit_plant.y ~ FALSE))
# #
```



**Phenology Calculations**  
 

```{r}
#Same as last year

BO_flowering <- BO_flr5 %>% 
  mutate(FFD = do.call(pmin, c(dplyr::select(., starts_with('FD_')), na.rm = TRUE)), #First flowering date
         days_sow2flower = as.numeric(FFD - Sow_Date), #days from lab sowing to flowering
         days_plant2flower = as.numeric(FFD - Plant_Date), #days from field planting to flowering
         LFD = do.call(pmax, c(dplyr::select(., starts_with('FD_')), na.rm = TRUE)), #last flowering date
         fl_duration = as.numeric(LFD - FFD)) %>%  #Flowering duration
  dplyr::select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, Sow_Date, Plant_Date, FFD, days_sow2flower, days_plant2flower, LFD, fl_duration, FD_01:FD_05, everything())



#TEMP FILE - delete later
write_csv(BO_flowering, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "BO_phenology_2023.csv"), na = "")

```


<br>  


#### Fruit and Seed counts  

Fruit & seed counts were recorded on the same spreadsheet, unlike last year. Loaded here and checked to make sure all ID info lines up with the Fitness data sheet.


```{r}
#Load Fruit & Seeds count data

BO_fruit_seeds <- read_csv(here::here("data_sheets", "2023_year2_data", "fruit_seed_counts", "BO_fruits_seeds_2023 - BO_fruit_seed_counts.csv"), skip = 1) %>% 
  rename(SegPos = Seg)


# #Remove rows where no fruits collected
# BO_fruit_seeds_2 <- BO_fruit_seeds %>% 
#   filter(total_fruits != 0)



##Check Fitness plants
BO_merge_test <- merge(BO_Fitness_2, BO_fruit_seeds, by = c("Transect", "SegPos", "Gen", "Recipient"), all.y = TRUE) %>% 
  mutate(FITTRUE = case_when(Fit_plant.x == Fit_plant.y ~ TRUE,
                             Fit_plant.x != Fit_plant.y ~ FALSE))


BO_fruit_seeds_2 <- merge(BO_Fitness_2, BO_fruit_seeds, by = c("Transect", "SegPos", "Gen", "Recipient", "Fit_plant"), all.x = TRUE) %>% 
  dplyr::select(-Sequence.y) %>% 
  rename(Sequence = Sequence.x) %>% 
  rename(Plant_ID = Fit_plant)
```

<br>  

#### Skeleton weight  

```{r}
# Check the column names of the skeleton data
skel_weight_BO <- read_csv(here::here("data_sheets", "2023_year2_data", "skeleton_data", "Skeleton Data 2023 - BO Skeletons.csv"))
colnames(skel_weight_BO)

# Rename columns using the actual names from colnames(skel_weight_BO)
skel_weight_BO <- read_csv(here::here("data_sheets", "2023_year2_data", "skeleton_data", "Skeleton Data 2023 - BO Skeletons.csv")) %>%  
  rename(
    position = side, 
    Gen = Gen,  # This stays the same since it's already "Gen"
    Fitness = `Fitness (y/n)`, 
    skel_dryweight_mg = Skeleton_Weights,  
    complete_incomplete = `Notes (complete or incomplete)`,  
    selected = `Chosen for Pulverization? (y/n)`
  )  %>% 
  arrange(Transect, Segment, position) %>% 
  mutate(SegPos = paste(Segment, position, sep = "")) %>% 
  dplyr::select(Transect, SegPos, Gen,  Recipient, Plant_ID, Fitness, skel_dryweight_mg) #complete/dplyr::selected not included


#old code
# skel_weight_BO <- read_csv(here::here("data_sheets", "2023_year2_data",  "skeleton_data", "Skeleton Data 2023 - BO Skeletons.csv")) %>% 
#   rename(position = side, Gen = 4, Fitness = 7, skel_dryweight_mg = 8, complete_incomplete = 9, dplyr::selected = 10) %>% 
#   arrange(Transect, Segment, position) %>% 
#   mutate(SegPos = paste(Segment, position, sep = "")) %>% 
#   dplyr::select(Transect, SegPos, Gen,  Recipient, Plant_ID, Fitness, skel_dryweight_mg) #complete/dplyr::selected not included

#create a dataframe with skeleton weight for each segment
skel_weight_BO_2 <- skel_weight_BO %>% 
  rename(skel_dryweight_mg_SEG = 7) %>% 
  dplyr::select(-Plant_ID)

#Check that IDs line up
skelcheck <- merge(BO_Fitness_2, skel_weight_BO_2, by = c("Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE)

```


**TEMP MERGED DATA FRAME**  
Merge trait data processed thus far
```{r}
#merging the skeleton data for each individual plant
BO_merged_1 <- merge(BO_flowering, skel_weight_BO, by = c("Transect", "SegPos", "Gen", "Recipient", "Plant_ID"), all.x = TRUE)  %>% dplyr::select(-c(Fitness, FD_01:FD_05, LeafD, FLR_P, Skeleton, Notes)) %>% arrange(Transect, Sequence, SegPos)

#merging the skeleton data for the entire segment
BO_merged_1.5 <- merge(BO_merged_1, skel_weight_BO_2, by = c("Transect", "SegPos", "Gen", "Recipient"), all.x = TRUE) %>% dplyr::select(-Fitness)%>% arrange(Transect, Sequence, SegPos)

####<-left off here
BO_merged_2 <- merge(BO_merged_1.5, BO_fruit_seeds_2, by = c("Sequence", "Transect", "SegPos", "Gen", "Recipient", "Donor", "Plant_ID"), all.x = TRUE) %>% dplyr::select(-c(N_CLOSED_fruit_D1:N_total_seeds_D6, Sym))  %>% arrange(Transect, Sequence, SegPos) %>% 
  dplyr::select(Sequence, Transect, SegPos, Gen, Donor, Recipient, Plant_ID, F_plant, everything(), -28)
```

<br>  


#### Seed weight  
[[OLD - UPDATE]]  

#### Seed weight  
[[UPDATED]]

```{r}

BO_seedmass_1 <- read_csv(here::here("data_sheets", "2023_year2_data", "seed_mass", "BO_Seed_Mass_2023.csv"), skip = 1) %>% 
  rename(SegPos = 3, Plant_ID = 7, mean_seed_mass_mg = 8) %>% 
  dplyr::select(Transect, SegPos, Gen,  Recipient, Plant_ID, mean_seed_mass_mg)
  
##Merge with fruit/seed count data frame

BO_merged_3 <- merge(BO_merged_2, BO_seedmass_1, by = c("Transect", "SegPos", "Recipient", "Plant_ID", "Gen"), all.x = TRUE)
#all.x means that any fitness plant x ID combos that don't match up won't be included.

```

<br>  

#### Corolla measurements  
[[UPDATED]]


```{r}
corollas_BO <- read_csv(here::here("data_sheets", "2023_year2_data", "corolla_leaf_data", "BO Corolla Diameters 2023.csv"), skip = 1) %>% 
  rename(Transect = 1, Recipient = 2, Gen = 3, Plant_ID = 5, SegPos = 6, corolla_diam_mm = 7, photo_date = 8) %>% 
  dplyr::select(Transect, Recipient, Gen, Plant_ID, SegPos, corolla_diam_mm, photo_date)

corollas_BO_1.5 <- corollas_BO %>% 
  filter(!is.na(corolla_diam_mm)) %>% 
  mutate(photo_date = as_date(photo_date)) %>% #correct missing dates manually
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% #Fix Recipient names
  mutate(Recipient = as.character(str_c("BO_", Recipient_num))) %>% 
  dplyr::select(-Recipient_num) %>% 
  arrange(Transect, Recipient, photo_date) %>% 
  distinct(Transect, Recipient, .keep_all = TRUE) #Keep only the earliest photo date data for duplicates

## Merge with main data 
BO_merged_4 <- merge(BO_merged_3, corollas_BO_1.5, by = c("Transect", "Recipient", "Gen", "SegPos", "Plant_ID"), all.x = TRUE)

#create a dataframe with corolla areas for each segment
corollas_BO_2 <- corollas_BO %>% 
  filter(!is.na(corolla_diam_mm)) %>% 
  mutate(photo_date = as_date(photo_date)) %>% #correct missing dates manually
  mutate(Recipient_num = sprintf("%03d", (parse_number(Recipient)))) %>% #Fix Recipient names
  mutate(Recipient = as.character(str_c("BO_", Recipient_num))) %>% 
  dplyr::select(-Recipient_num, -Plant_ID) %>% 
  arrange(Transect, Recipient, photo_date) %>% 
  distinct(Transect, Recipient, .keep_all = TRUE) %>% 
  rename (corolla_diam_mm_SEG = 5, photo_date_SEG = 6) #Keep only the earliest photo date data for duplicates


BO_merged_4.5 <- merge(BO_merged_4, corollas_BO_2, by = c("Transect", "Recipient", "Gen", "SegPos"), all.x = TRUE)
```


<br>  


#### Leaf Area & Weight  
[[UPDATED]]


```{r}
leaves_BO <- read_csv(here::here("data_sheets", "2023_year2_data", "corolla_leaf_data", "BO_Leaf_Area_Weight_2023.csv"), skip = 1) %>% 
  rename(Transect = 1, Gen = 2, Donor= 3, Recipient = 4, SegPos = 5, Plant_ID = 6, leaf_area_mm2 = 7, leaf_dryweight_mg = 10) %>% 
  dplyr::select(Transect, Recipient, Gen, Donor, SegPos, Plant_ID, leaf_area_mm2, leaf_dryweight_mg) %>%
  filter(!is.na(leaf_area_mm2)) %>% #exlcude NA leaf areas?
  mutate(leaf_area_mm2 = as.numeric(leaf_area_mm2))


#Chec for duplicates
leaves_BO_dup_check <- leaves_BO %>% 
  mutate(duped1 = duplicated(dplyr::select(leaves_BO, c(Transect, Recipient, Gen, Donor, SegPos))),
         duped2 = duplicated(dplyr::select(leaves_BO, c(Transect, Recipient, Gen, Donor, SegPos)), fromLast = TRUE),
         duped = case_when(duped1|duped2 == TRUE ~ TRUE,
                           TRUE ~ FALSE)) %>% #highlight all duplicates
  dplyr::select(-c(duped1, duped2))
# No "TRUE"s found, so no duplicates, no need to remove them

#Merge in to main data 
BO_merged_5 <- merge(BO_merged_4.5, leaves_BO, by = c("Transect", "Recipient", "Donor", "Gen", "SegPos", "Plant_ID"), all.x = TRUE) #replicates leaf area & weight for every plant of a transect recipient combo

#create a dataframe with skeleton weight for each segment
leaves_BO_2 <- leaves_BO %>%
  rename(leaf_area_mm2_SEG = 7, leaf_dryweight_mg_SEG = 8) %>%
  dplyr::select(-Plant_ID)

BO_merged_5.5 <- merge(BO_merged_5, leaves_BO_2, by = c("Transect", "Recipient", "Donor", "Gen", "SegPos"), all.x = TRUE) #replicates leaf area & weight for every plant of a transect recipient combo

#Create a column called SLA in the merged dataframe for both individual plant and segement
BO_merged_6 <- BO_merged_5.5 %>%
  mutate(SLA_SEG = leaf_area_mm2_SEG / leaf_dryweight_mg_SEG) %>%
  mutate(SLA = leaf_area_mm2 / leaf_dryweight_mg) %>%
  mutate(LMA_SEG = 1 / SLA_SEG) %>%
  mutate(LMA = 1 / SLA)

```

<br>  


#### d13C data  
[[UPDATED]]

```{r}
BO_d13C_2023 <- read_csv(here::here("data_sheets", "2023_year2_data", "isotope_data", "BO_d13C_2023.csv")) %>% 
  dplyr::select(1,2,2,3,5,6,9) %>% 
  rename(sample_ID = 1, d13C = 2, totalC_ug = 3, sample_weight_mg = 6)
#Save as UTF-8 file from excel to avoid import error
  
BO_d13C_2023_1.5 <- BO_d13C_2023 %>%
  dplyr::select(sample_ID, d13C, sample_weight_mg)

BO_d13C_2023_2 <- BO_d13C_2023 %>%
  mutate(sample_ID = substr(sample_ID, 1, nchar(sample_ID) - 2)) %>%
  dplyr::select(sample_ID, d13C, sample_weight_mg) %>%
  rename(sample_ID_SEG = 1, d13C_SEG = 2, sample_weight_mg_SEG = 3)

BO_merged_6.5 <- BO_merged_6 %>%
  mutate(sample_ID_SEG = paste(Recipient, Gen, str_pad(Transect, width = 3, pad = "0"), SegPos, sep = "-")) %>%
  mutate(sample_ID_SEG = str_replace(sample_ID_SEG, "^([^_]*)_", "\\1-"))

BO_merged_7 <- BO_merged_6.5 %>%
  mutate(sample_ID = paste(Recipient, Gen, str_pad(Transect, width = 3, pad = "0"), SegPos, Plant_ID, sep = "-")) %>%
  mutate(sample_ID = str_replace(sample_ID, "^([^_]*)_", "\\1-"))

#merge a dataframe with isotope data for each plant
BO_merged_7.5 <- merge(BO_merged_7, BO_d13C_2023_1.5, by = c("sample_ID"), all.x = TRUE)

#merge a dataframe with isotope data for each segment
BO_merged_8 <- merge(BO_merged_7.5, BO_d13C_2023_2, by = "sample_ID_SEG", all.x = TRUE)
#Merge by Recipient and Transect only: genotype-wide estimates of d13C since some plants were NF

# Subset to exclude rows where d13C is NA
testdf <- BO_merged_8 %>% 
  filter(F_plant == "TRUE") %>%
  filter(!is.na(d13C)) # 1 plant is a NF plant, 1 NF + 187 F plants = 188 observations

isotope_comparison <- anti_join(BO_d13C_2023, BO_merged_8, by = c("sample_ID", "d13C", "sample_weight_mg")) %>%
  dplyr::select(d13C, sample_ID)

```


<br>  


**Error Checks**  

**Error checks**  

Checking a variety of IDs and trait data to make sure everything lines up.  

```{r}

#Chec for duplicates
BO_dup_check <- BO_merged_8 %>% 
  mutate(duped1 = duplicated(dplyr::select(BO_merged_8, c(sample_ID))),
         duped2 = duplicated(dplyr::select(BO_merged_8, c(sample_ID)), fromLast = TRUE),
         duped = case_when(duped1|duped2 == TRUE ~ TRUE,
                           TRUE ~ FALSE)) %>% #highlight all duplicates
  dplyr::select(-c(duped1, duped2)) %>%
  filter(duped == "TRUE")
# No "TRUE"s found, so no duplicates, no need to remove them

##Check for mismatched (fitness) plants

BO_Echeck1 <- BO_merged_8 %>% 
  filter(F_plant == FALSE & total_fruits != 0) #do the same for seed count, using `>` and `!=`

BO_Echeck2 <- BO_merged_8 %>% 
  filter(F_plant == TRUE) %>% 
  filter(ALL_closed_fruit == 0 & total_seeds > 0)

##NA/NaN/Info values for mean seeds/est seed prod
BO_Echeck3 <- BO_merged_8 %>% 
  filter(F_plant == TRUE & is.na(as.numeric(avg_seeds_per_fruit)))

BO_Echeck4 <- BO_merged_8 %>% 
  filter(F_plant == TRUE & !is.finite(avg_seeds_per_fruit))


## Check for NA seed counts for fitness plants (when fruit production > 0)
#Will be empty when seed_ct NAs are replaced with 0
BO_Echeck5 <- BO_merged_8 %>% 
  filter(F_plant == TRUE & is.na(total_seeds))
##NAs present here before setting seed_ct NAs to 0 were from fitness plants that had 0 closed fruit


#Seed mass
BO_Echeck6 <- BO_merged_8 %>% 
  filter(F_plant == FALSE & !is.na(mean_seed_mass_mg))


```

<br>  


### Trait calculations 

Some trait calculations performed in `Fitness_and_EDA.Rmd.  

Fruits were only collected from fitness plants, so only survival to flowering is calcualted for all plants (fitness and non-fitness).

UPDATE with latest merged df...

```{r}

#Survival measures
#Added var for closed fruit produced
#renamed "survival" to seed production
#
BO_merged_9 <- BO_merged_8 %>% 
  mutate(surv_to_flower = case_when(!is.na(FFD) ~ 1, #plants that survived to flower
                                  is.na(FFD) ~ 0),
         surv_to_fruitprod = case_when(F_plant == TRUE & total_fruits > 0 ~ 1, #plants that survived to produce fruit
                                       F_plant == TRUE & total_fruits == 0 ~ 0),
         prod_closed_fruits = case_when(F_plant == TRUE & ALL_closed_fruit > 0 ~ 1,
                                            F_plant == TRUE & ALL_closed_fruit == 0 ~ 0),
         prod_filled_seeds = case_when(F_plant == TRUE & ALL_filled_seeds > 0 ~ 1, #plants produced filled seeds
                                      F_plant == TRUE & ALL_filled_seeds == 0 ~ 0)) #else = 0

#NOTE
#Adjusting fecundity/fitness:
#The issue of how to handle 0s vs missing data in order to accurately calculate fitness has led us to apply the following corrections:
# Estimated fecundity (seeds per fruit x total fruits) should be set to *0* when closed fruits >= 1 & filled seeds == 0; when closed fruits == open fruits == 0
#Set fecundity to NA when closed = 0 & open >= 1

#Mean seeds per fruit left as-is (NA when no closed fruits collected)
         

BO_merged_10 <- BO_merged_9 %>% 
  rename(mean_seeds_per_fruit = avg_seeds_per_fruit) %>%
  mutate(across(c(est_fecundity, mean_seeds_per_fruit), ~ as.numeric(na_if(., NaN)))) %>% 
#Converts to NA if any of the values are "NaN"
  mutate(mean_seeds_per_fruit = case_when(
          F_plant == TRUE & ALL_closed_fruit == 0 ~ NA, # NA when 0 closed fruits
          TRUE ~ mean_seeds_per_fruit)) %>%  #else, original value
          mutate(est_fecundity = case_when(
          F_plant == TRUE & ALL_closed_fruit == 0 & ALL_open_fruit == 0 ~ 0,
          F_plant == TRUE & ALL_closed_fruit == 0 & ALL_open_fruit >= 1 ~ NA,
          F_plant == TRUE & !is.na(est_fecundity) ~ est_fecundity))


```

<br>  



### Master sheets (2023)  

Full and summarized data sheets for all data collected in the 2022-2023 field season. 
These are merged with the year 1 data sheets elsewhere.  

```{r}
## 'Long' version with all data - one row per plant ID
#Sort by transect and ID variables, then by Fitness plants and those that flowered...

######
######

BO_merged_full_2023 <- BO_merged_10 %>% 
  filter(F_plant == TRUE)
##NOTES - there are still fitness plants that never flowered... keep as 0s
#fitness plants only

write_csv(BO_merged_full_2023, here::here("data_sheets", "2023_year2_data", "compiled_sheets_2023", "BO_merged_full_2023.csv"), na = "")


######
######

BO_mastersheet_v1_2023 <- BO_merged_10 %>% 
  mutate(Year = 2023) %>% 
  dplyr::select(Year, Gen, Transect, Sequence, Donor, Recipient, Plant_ID, F_plant, everything()) %>% 
  arrange(Transect, Sequence, Recipient, Plant_ID) %>%
  rename(open_fruits = ALL_open_fruit, closed_fruits = ALL_closed_fruit, filled_seeds = ALL_filled_seeds, msm_all = mean_seed_mass_mg)

write_csv(BO_mastersheet_v1_2023, here::here("data_sheets", "compiled_sheets", "BO_mastersheet_full_2023.csv"))
#Renamed to full - has records of all plants (including non-fitness)

#
## Fitness Plants Only
#

BO_mastersheet_F_mains_2023 <-  BO_mastersheet_v1_2023 %>% 
  filter(F_plant == TRUE) #Only main fitness plants

##Note: remaining NA values for fitness measures (mean seeds per fruit; est seed production) are the results of 0 closed fruit counts. Treat these as 0 fitness or as NAs to remove???


write_csv(BO_mastersheet_F_mains_2023, here::here("data_sheets", "compiled_sheets", "BO_mastersheet_Fitness-mains_2023.csv"))

```



